\chapter{Abilities}

\begin{multicols}{2}

Skills can only be acquired once, unless stated otherwise in the description. When a technique is acquired multiple times, it receives an I, II, III next to its name, indicating how many times it has been chosen.

A character does not need to be at level x to gain the XP from level x. Some techniques only require XP at level x, but can be learned even if a character is at a lower proficiency level.

It is not possible to acquire a technique unless the requirements are met.



\section{DEX}

\subsection{Athletics}

Athletics represents the character's movement abilities. 

% \abil{Combo}{passive}{6 XP, DEX 3/6}{-}{ Increases the amount of AP granted by the fighting surge by 1/2 AP.}

\abil{Sprinter I/II}{passive}{6 XP, CON 3/5}{}{ I. Increases running speed by 1/2m.  }

\abil{Jumper}{passive}{6 XP, CON 3/5}{}{ I. Increases jump distance by 1/2m.  }

\abil{Climber I/II/III}{passive}{automatic in climbing 1/3/5}{}{+2/4/6 bonus for climbing. I. Learns to spend 1 STA on a critical to double climbing speed. II. Increases crawling and climbing speed to 0.5m xMM per AP. III. The character is not considered prone while climbing. }

\abil{Gymnast I/II/III}{passive}{automatic in balance 1/3/5}{ }{ I. Balance tests can be performed safely. II. Falling damage is halved from the original damage when landing on feet. III. Falling damage is reduced to one third if landing on feet. }

\abil{Swimmer I/II/III}{passive}{6XP, automatic in swimming 1/3/5}{}{ I. Can perform safe swimming tests and swim quickly. II. Spends 1 STA for 4 AP of fast swimming or 10 AP on criticals. III. Increases normal swimming speed to 0.5m xMM and fast swimming to 1m xMM. }

\abil{Riding}{passive}{4 XP, DEX 1/3/5}{}{ I. Learns to give commands to walk and run for mounts capable of understanding the command. II. The rider can utilize the mount's jumping, standing, swimming, and climbing movements if it is capable of understanding the commands. III. The rider can utilize the flight of their mount while riding.}




% \abil{Power Attack}{2 STA}{4 XP, Athletics 4}{Melee 3}{ Use muscle overcoming to make an attack stronger, increasing the STR attribute in the damage calculation, in the hook or supported attack test, or in the strength requirement of the attack. The strength bonus is equal to that of muscle overcoming.}

% \abil{Vascular Control }{passive}{6 XP, Athletics 3}{}{ Can spend all AP during choking without losing STA, unless left with negative AP. }

% \abil { Conditioning I/II/III } {passive}{6 XP, Athletics 2/4/6 } { STR 12/14/16 } { I. Reduces penalties for skill checks when wearing armor by 1/1/2. II. Also reduces the STA penalty for wearing armor by 1 and removes the exhaustion penalty for using heavy armor. III. Reduces the AP penalty for wearing armor by 1 and removes the difficulty to stand up. }

% \abil{Runner I/II}{passive}{6 XP, Athletics 2/5}{}{ I. Increases running speed by +1m/2m. II. Increases jump height while running by +1m. }

% \abil{Power Run}{1 STA}{6 XP, Athletics 4}{}{ Increases running speed by 3m and long jump by +2m, but these actions cost 1 STA more.}

% \abil { Tough I/II/III } {passive}{6 XP, Athletics 2/4/6 } { STR 12/14/16 } { Increases natural RES by 1/2/3.  }


% \abil { Jumper I/II/III} {passive}{ 6 XP, Athletics 1/3/5 } { AGI 12/15/18 } { A jump becomes longer by 1/2/3m and higher by 0/1/1m. }

% \abil{Energy Saturation I/II/III}{passive}{8 XP, Athletics 2/4/6}{}{ Learns to saturate the body with energy, increasing the learned CON by +1/2/3. }

% \inna{Muscle Overcoming}{2 STA}{6 XP, Athletics 4}{Athletics 3}{self}{instantaneous}{Spend 2 STA to increase Strength in a strength test, climbing, or grabbing by STR/2 for a single action or reaction. }



% end -----------------------------------------------------------------------

\subsection{Ranged Combat}

% \abil{Adrenaline}{passive}{4 XP, DEX 2}{}{ The character can spend 1 extra STA to move the equivalent of one more AP of movement after the explosion or projectile.}

\abil{Mounted Shooter}{passive}{4 XP, DEX 3/5}{Archer, Rider I}{ I. Reduces the penalty for shooting while riding by 1/2. }

\abil{Thrower I/II}{passive}{6 XP, DEX 3/5}{}{The range of thrown weapons is increased by 50/100\%. II. Reduces the distance penalty from -2 to -1.}

\abil{Archer}{passive}{6 XP, DEX 1}{}{Learns to shoot with bows, removing the penalty, and allows for quick and precise shots.}

\abil{Elite Sniper }{passive}{6XP, DEX 5}{Archer}{  Precise shots cost 1 AP less.}

\abil{Quick Shooter I/II/III}{passive}{6XP, DEX 2/4/6}{Archer}{ I. Reduces the penalty for quick shots to -2/-1/0. }

\abil{Strafe}{passive}{6XP, DEX 5}{Quick Shooter II}{ I. Shoots a short bow 3x within a single fighting surge, but takes -3 penalty to accuracy. }

\abil{Crossbow and Shield}{passive}{4 XP, DEX 2}{}{ Allows the character to be in guard with a tower shield while shooting and reloading a crossbow or arbalest.}

\abil{Steady Arms}{passive}{6 XP, DEX 3}{}{Can move the equivalent of 1 AP of simple movement while shooting, without additional AP cost. It is considered that the shot came from the shooter's final position. Opportunity attacks are triggered at all positions along the path.}

\abil{Archer's Muscles I/II}{passive}{6 XP, CON 2/4}{Archer}{ Provides a strength bonus of +1/2 to meet the strength requirement of bows.}

\abil{Archer's Reflexes}{passive}{6XP, DEX 3}{Quick Shooter I}{Allows for a quick shot at any visible target in reaction to any action of the target that triggers an opportunity attack and the shooter has perceived.}

\abil{Precise Reflexes}{passive}{6 XP, DEX 4}{DEX 4}{ If the character uses reflexes without moving, reflexes only cost 1 AP.}

\abil{Quick Reload I/II }{passive}{6 XP, DEX 2/4}{}{Reloads weapons 1/2 AP faster.}


% \abil{Cadenced Shot}{passive}{4 XP, ranged combat 3}{}{ Normal shots and precise shots can exchange the STA cost for AP. }

\abil{Jump Shot}{Attack cost + +1STA}{6 XP, DEX 4}{}{ Spends 1 extra STA to jump backward and perform a quick shot or throw, avoiding opportunity attacks if the jump takes them out of range.}

% \abil{Tactical Movement}{passive}{6 XP, Ranged Combat 2}{}{ Whenever the character's last action was running, they receive a +3 bonus to SD against subsequent ranged attacks.}


% end -----------------------------------------------------------------------

\subsection{Melee Combat}

% \abil{Combat Acrobatics}{passive}{6 XP, melee combat 3}{AGI 15}{The evasive jump grants +4 extra dodge instead of +3.}

% \abil{Grapple in Battle I/II/III}{passive}{6 XP, melee combat 2/4/6}{}{Reduces the penalties for being grappled from 3 to 2/1/0.}

\abil{Blade Alignment }{passive}{6 XP, DEX 3}{}{ Increases the RES of blade weapon strikes, that is, with PEN = 1x, by 5xDM.}

\abil{Gentle Art }{passive}{6 XP, DEX 4}{}{ You can take a -2 penalty on grapple tests in exchange for reducing the STA cost by 1.}

\abil{ Clinch }{}{6 XP, DEX 2}{-}{ You can initiate a grapple without cost whenever you redirect an attack in melee range and the attack misses. The character must have the grapple property on any of their equipped weapons. Hands only count if they are free.}

\abil{Grappling Combat }{sustained}{6 XP, DEX 2}{-}{ Grants attacks for grappling that cost 4AP+1STA and inflict wounds directly. Critical hits can cause special effects: Suffocate the enemy, break an arm, and break a leg. Hits immobilize the opponent.}

\abil{Counterattack}{AP of the attack}{6 XP, DEX 2}{}{ You can attack in response to an attack, as long as you are within range. Both attacks are made against the opponent's SD. The attack with the higher result hits first, having the chance to interrupt the opponent. The counterattack receives -2 to hit.}

\abil{Battle Coordination }{passive}{4 XP, DEX 2}{}{ When there are two allies within melee range of the same enemy, the enemy's attacks against one ally trigger opportunity attacks from the other, provided they have this ability. The opportunity attack occurs after what triggered it. Wide attacks do not trigger this, since it is made in an area.}

\abil{Defender}{passive}{6 XP, DEX 3}{}{ Allows you to spend 1 STA to move 1 normal movement as a reaction to position yourself to defend someone.}

\abil{Offensive Defense}{passive}{6 XP, DEX 2}{}{ When the opponent grazes or misses against the redirect, you can move one basic movement forward without additional AP cost. }
% \abil{Weapon Specialization I/II/III}{passive}{10 XP, melee combat 3/5/6}{}{ Choose to receive +1/2/3 to hit, block, evade, and grapple while using a specific type of weapon. You can learn this ability for each weapon.}

\abil{Hook }{passive}{6 XP, DEX 3}{}{ Increases the threat area of the reaction, allowing you to spend 1 STA to move 1 basic movement before using the hook attack as a reaction. }

% \abil{Hook }{passive}{6 XP, DEX 3}{}{ Increases the impact of knocking down and disarming by +0.5x STR, this impact does not influence the PEN calculation. }

\abil{Death Spin}{passive}{6 XP, DEX 2}{}{Sweeping attacks with two-handed weapons can cover up to 360 degrees.}

% \abil{Combat Instincts I/II/III}{passive}{8 XP, DEX 2/4/6}{}{ Increases armor coverage by 1/2/3 if not immobilized.}

\abil{Charge}{passive}{6 XP, DEX 1}{}{ Does not need to stop running to perform a melee attack and can attack any character in range during any step in movement trajectory. Can only attack one target per running block and must spend AP accordingly.}

\abil{Lancer }{passive}{6 XP, DEX 4}{Rider I}{ Removes the penalty from braced attack while mounted. }

% \abil{ Lumberjack I/II}{passive}{6 XP, DEX 3/5}{}{ Reduces the penalty for hitting heavy attacks by 1/2.  }
% \abil{ Lumberjack I/II}{passive}{6 XP, DEX 3/5}{}{ I. Increases the base damage of attacks with heavy property I by +STR/6 and heavy II or more by STR/5. II. Increases to STR/5 and STR/4. }

\abil{Technical Lift}{passive}{4 XP, DEX 1}{}{ The character can spend 1 STA to perform an evasive jump and stand up in the same action. The penalty for being down still applies during this movement. }

% \abil{Shield Marking}{passive}{6 XP, DEX 2}{}{ Maintains contact of the shield with the opponent's weapon, marking them and increasing the AP cost of their attacks against your shield by 1 after successfully blocking an attack (receiving a miss or graze). Only one opponent can be marked at a time. Blocking another opponent, receiving a hit or critical from the marked opponent, or leaving their threat area removes the marking. }

% \abil{Master with Blades}{passive}{10 XP, DEX 2/4/6}{}{ Increases the PEN of melee weapons with PEN 2x or greater by 2/4/6.}

\abil{Lunge}{passive}{6 XP, DEX 1}{}{ Allows you to move one basic movement for 1 STA before using an attack as an action. This movement triggers opportunity attacks and braced attacks normally.}

\abil{Shield Wall}{passive}{6 XP, DEX 1 }{}{Gains +2 to block if holding a shield and adjacent to another character with a shield.}

\abil{Pikeman }{passive}{6 XP, DEX 3}{}{ Increases the threat area of the reaction, allowing you to spend 1 STA to move 1 AP before using the braced attack as a reaction. }

% \abil{ Retaliation }{1AP}{6 XP, DEX 3}{}{ Spend 1 AP to block with armor and move 1xMM. Gain an opportunity attack as long as you are not interrupted.}

\abil{Riposte }{AP of the attack-1}{8 XP, DEX 3}{}{After redirecting a melee attack without being hit or critted, the character can make an opportunity attack as a reaction. The attack costs 1 AP less than the normal attack. The attack costs 2 AP less if made with an object different from the one used for defense. }

% \abil{Subjugate I/II}{passive}{6 XP, melee combat 3/5}{}{Increases the bonus of precise attacks to hit vulnerable attacks by +1/2. }

\section{SPI}

\newcommand{\convic}[2]{
    \textbf{ Level #1}  \\
     #2 \\
  }  

\subsection{Conviction of Adaptation}
\textbf{Worldview - Chaos - Introspective }

They believe that the universe is chaos and that those who do not adapt should be left behind. 

\textbf{Promoted Attitudes:} Learns to take any test in a risky manner. Gains 1XP each time they roll a 9 or 10 on a risky test. \\
\textbf{Prohibited Attitudes:} Loses 1 XP each time they fail or graze by on a test that is not risky. \\

 \convic{1}{ Can use adaptation XP as any type of XP considering the character's talent +1. When doing this, the XP also increases the progress of Adaptation.  }
 
 \subsection{ Conviction of Domination }
 \textbf{Worldview - Chaos - Extrospective }
 
 They believe that most people are mediocre and incapable of acting in their own benefit. This justifies their control over others.
 
 \textbf{Promoted Attitudes:} Gains 1 XP each time a social action or the effect of affliction or compulsion hits or crits. At most, once per turn.
\textbf{Prohibited Attitudes:} Loses 1 XP each time a social action, mental affliction or compulsion grazes by or fails. At most, once per turn.

\convic{1}{ Receives +1 in Deception and Persuasion.}

\convic{2}{ Can spend 2 AP as a reaction to a player's limit action to give them orders, which works like the suggestion spell. Do the same with one character of your choice if it reaches your limit action of Abuser. }

\convic{3}{ Any time you crit in any action against someone in combat or when they miss in one, you can mock them at the same time for 1 extra AP, which works as provocation and gets +2 to the provocation test.}

\convic{4}{Whenever an opponent uses suggestion or command and get a miss, they must make a will test against your DL for psychic spells to avoid losing 2 AP. }

\convic{5}{ Gets +1 to all tests against characters affected by any mind altering status effects. Also increases the DL of any spells used against such characters.}

\subsection{ Conviction of Stoicism }
\textbf{Worldview - Order - Introspective }

\textbf{Promoted Attitudes:} Gains 1 XP each time a prepared action hits or crits.
\textbf{Prohibited Attitudes:} Loses 1 XP each time they graze or miss in a prepared action.

\convic{1/3}{ Prepared actions gain +1/2 on the test of the prepared action. }

\convic{2}{Can learn with conviction XP everything in the section on Cunning. When doing so, the XP also increases the progress of Stoicism.}

\convic{2/4}{ Receives +1/2 to the DL of the prepared spell.}

\convic{3}{Can prepare 2 actions with 1 PA.}

\convic{4}{ Whenever you win a cunning test to act before another character, you gain +1 in all skill tests against that character.}

\convic{5}{Receives +1 AP for each crit in the prepared action.}

\subsection{ Conviction of Fatalism }
\textbf{Temperament - Order - Introspective }

They believe that there is a natural order. Things should be as they are, and it is wrong to fight against them. Their purpose is to fulfill their role.

\textbf{Promoted Attitudes:} Gains 1XP each time they overcome a will test with a DL greater than or equal to the character's will or 2XP if the DL is 3 or more higher.
\textbf{Prohibited Attitudes:} Loses 1 XP each time they obey a compulsion or for each turn they are affected by a mental affliction.

\convic{1}{ Active will tests can be made with 2 AP.}

\convic{2/4}{ Receives +1/2 in Will. Still triggers the XP gain when the will test was overcome by 1/2. }

\convic{2/4}{ I. The character ignores the effect of tiredness II. Ignores the effect of exhaustion. }

\convic{3}{The character gains +1 to all skills in any turn when the morale test has a DL higer than 5.}

\convic{5}{ Character take half penalty for mental afflictions.} 

% Rage
\subsection{Conviction of Ferocity}
\textbf{Temperament - Chaos - Extrospective }

They understand that everything can be resolved with more strength. 

\textbf{Promoted Attitudes:} Gains 1 XP of conviction each time they start a turn with 0 AP or less. They can select one opponent at a time as a target when they are angry to gain 1 XP if they incapacitate them on the same turn.
\textbf{ Prohibited Attitudes:} Loses 3 XP for surrendering or fleeing from an ongoing combat. Loses 1 XP for each turn that ends in fear.

\convic{1}{The mental penalty of rage is removed. Every time the character uses combat surge while enraged, the morale test DL for all oponents in a 5m range at the moment of the surge increases by the character's ferocity level. This does not stack with more instances of this effect.}

\convic{2}{ Learns to enter a rage voluntarily with an active stress test DL 3. If the character is afraid, the DL is the same as what caused their fear. }

\convic{3/5}{The character cannot be interrupted by receiving tier 2 injuries while in rage. II. They are also not interrupted by receiving tier 3 injuries. Does not fall unconscious from IL when enraged and only dies at 6x IL. }

\convic{4}{The character ignores the effect of exhaustion and sickness during rage.}

\convic{4}{The character learns spiritual strike and martial spirit. Uses will instead of animancy against the opponent. }

\subsection{ Guardian Conviction }
\textbf{Temperament - Order - Extrospective }

They believe that what exists should not be destroyed without reason.

\textbf{Promoted Attitudes:} Gains 1 XP each time they pass the test to protect someone.  
\textbf{Prohibited Attitudes:} Loses 2 XP when refusing to protect.

% Minala
\convic{1/3/5}{When they realize that an ally has been attacked, they can make a protection surge as a reaction. The surge costs 3 STA and grants them 6/7/8 AP to immediately assist the ally. The character's actions must necessarily heal the ally or be defensive reactions.}

\convic{3}{ Every ally gets +2 in their morale check any turn the guardian uses its protection surge.}

\convic{2/4}{Gains +1/2 to block when making the protection surge to protect someone. Also gains +1/2 in sorcery for any healing action.}

\convic{5}{Gains +1 to all skills when making the surge to protect someone.}

%revisar
\subsection{Hedonism Conviction}
\textbf{Temperament - Chaos - Introspective }

They believe that life is meant to pursue good things and pleasure. They learn to seek happiness in all things and to enjoy them in the best way possible, without letting them destroy them. They can consume hedonism karma to alleviate the negative effects of vices.

\textbf{Promoted Attitudes:} Enjoying their vices.  
\textbf{Prohibited Attitudes:} None.

%Hedonism
\convic{1/3/5}{The character has +1/2/3 to will in any test that forces them to deny a vice.}

\convic{2/4}{The mental penalty for being intoxicated decreases by 1/2. At level 2, the character ignores the effect of tiredness when intoxicated.}

\convic{2/4}{Gains +1/2 to the persuasion test if they have the vice "center of attention."}

\convic{2/4}{Receives +1/2 for prestidigitation if they have the vice "Kleptomaniac."}

\convic{2/4}{Receives +1/2 for seduction tests if they have the vice "Nymphomaniac."}

\convic{2/4}{Receives +1/2 for taunting if they have the limit action of "Abuser."}

% \convic{2/4}{Receives +1/2 for persuasion tests if they have the limit action of "Traitor."}

\convic{3}{Gains double XP in informal knowledge if the activity they engage in involves fun.}

\convic{3}{Gives +Charisma/2 to the team's morale tests if they say something funny. This can be done after another character fails the test.}

\convic{3}{Gains the ability to roll exploded die twice and pick the most extreme option.}

% \convic{3}{+1 XP for each book read. Learning a spell from the grimoire counts as a book.}

\subsection{ Providentialism Conviction }
\textbf{Worldview - Order - Extrospective }

They believe and find comfort in the belief in a good universal order. If everyone acted correctly, according to the order, everything would work well.

\textbf{Promoted Attitudes:} Follow the moral code described in the skills.  
\textbf{Prohibited Attitudes:} The character loses 2 karma each time they lie. Loses 3 karma for allowing a creature with INT greater than or equal to 5 to be killed without granting the right to a trial. Self-defense or protection of the innocent does not count. Loses 6 karma if they unjustly kill with their own hands.

\convic{1}{Whenever they are about to kill a creature considered intelligent, they must grant it the right to a trial. If it truly agrees to follow the moral code, the character is prohibited from killing it, mutilating it, putting it in life-threatening situations, or allowing others to do so. Gains 2 XP each time they spare someone or a group of people.}

\convic{1}{Whenever they make a relevant promise and manage to fulfill it, they gain at least 1 XP. They can earn more for larger promises.}

\convic{1}{Whenever the character or someone on their team is lying, make a Will test DL 5. If they succeed or crit, they must rectify the lie (this does not mean they are obliged to expose the truth). If they do not lie, regardless of passing the test, they gain 1 XP.}

\convic{2}{Gains the ability to make any test safely and preferring numbers closer to 6 instead of 5 if the test is in line to complete a promise.}

\convic{3}{Gains +2 in social actions and mind altering spells whenever they are judging surrendered opponents.}

\convic{4}{Gives +Charisma/2 to the morale tests of the entire team when they say something encouraging. This can be done after another character fails the test.}

\convic{5}{When the character is trying to fulfill a promise and receives a wound, they can immediately perform an action surge, even if they have no STA.}

% end -----------------------------------------------------------------------

\subsection{Deities and Devotion}

\abil{Cleric }{passive}{Automatic at Devotion 1}{Devotion}{ The character becomes a cleric of a deity. It is only possible to become a cleric if the character is not devoted to another and shares a common conviction with the deity. This allows all abilities from the deity's list to be learned with Devotion XP. The Devotion level + SPI is used as a skill to perform these spells.}

% \abil{ Divine Guide }{passive}{4 XP, Devotion 3}{Cleric}{ Can cast a careful spell with only 1 additional AP for spells associated with the deity. }

\abil{Miracle}{passive}{automatic at Devotion 3}{Cleric}{ Allows the casting of any spell contained within the deity's list, even without knowing it or having sufficient level, but incurs a -2 penalty on the spellcasting test for these and must perform them in a risky manner.  }

\abil{Prayer}{passive}{automatic at Devotion 2}{Cleric}{ Improves the resting action. Focuses on the divine energy around, increasing STA regeneration by Devotion/2. Recovers 100\% of Devotion if in a temple or sacred place. Cannot be used in a location outside divine influence.}

% \abil{Divine Overload}{passive}{6 XP, Devotion 4}{Cleric}{ Overloading a spell from the deity's list increases the stamina cost by 50\% instead of 100\%.}


% \abil{Divine Intervention}{passive}{6 XP, Conviction 4}{Divine Spellcasting}{ Once per day, the character has a 50\% chance to cast any spell or ability they know without any AP or stamina cost when their HP drops below 0.}

\subsection{Vices and Traumas:}

\textbf{How to acquire:} the character can take a vice voluntarily and gain 6 karma  at character creation. When players fail a mission or something very bad happens, the GM may allow players to get the karma reward as long as they take on a vice.

It is possible to have a maximum of 4 vices. Overcoming vices costs 6 karma. So vices are like shortcuts or borrowed advancement.

\abil{Alcoholic}{passive}{ }{ }{ Penalty of -3 for moral tests if not intoxicated, bonus of +1 if intoxicated. Gains 1/2/3 Hedonism XP for entering combat with alcoholism I/II or III, respectively.}

\abil{Center of Attention}{passive}{ }{Vice}{+1 Hedonism XP for each time successfully distracting or provoking someone and -1 if failing.}

\abil{Kleptomania}{passive}{ }{}{ May be compelled to steal something. Must make a will test with a difficulty of 5 to avoid being compelled. Loses 1 XP if not stealing and gains 1 Hedonism XP if attempting to steal. }

\abil{Explosive}{passive}{ }{Trauma}{ Receives -3 on any test to resist becoming angry or to calm down when already furious.}

\abil{Phobia of x}{passive}{ }{}{ Has a moral penalty when in the presence of the phobia's object. Gains 1 Ferocity XP if incapacitating the target of fear. }

\abil{Fear of Heights}{passive}{ }{}{ Receives -5 for moral tests when in high places, with a risk of falling.}

\abil{Nymphomaniac}{passive}{}{}{ Receives -3 to resist seduction tests. Gains 1 Hedonism XP each time they successfully seduce someone, but loses 1 XP whenever they fail. }

\abil{Hungry}{passive}{}{}{Receives -3 mental if suffering any distress from hunger.}

\abil{Insomnia}{passive}{}{}{Receives -2 to health when trying to rest, but gains +2 to cunning while unconscious.}

% end --------------------------------------------------------------------------------------------------

\section{INT}

% Basic

\abil{Literacy}{passive}{ 6 XP INT 1}{Language 3}{Learns to read and write.}

%Art
\subsection{Art}

\abil{Performance I/II/III}{passive}{4 XP, INT 2/4/6}{}{I. Allows the character to know a form of art, such as playing an instrument, magic tricks, jokes, etc. Gains +1/2/3 to tests to distract the audience.  }

\subsection{Animancy}

% \abil{Psychology I/II/III}{passive}{8 XP, INT 2/4/6}{}{I. Increases by +1/2/3 to your knowledge when trying to resist a social action. Allows for more elaborate opinions on the behavior of creatures in general. }

% Medicine and botany / Biomancy

\subsection{Biomancy}

\abil{First Aid}{ritual}{automatic at Biomancy 1}{}{Learn to heal wounds, make splints, and fix broken bones. By spending an exploration turn, you can heal 2 IL to an injured creature if used within 4 hours after it has been hurt. }

% Tools

\subsection{Engineering and Alchemy}


\abil{Armorer }{passive}{automatic at Armorer 1}{}{ I. Allows the character to make simple repairs and maintain weapons and armor.}

\abil{Mechanic }{passive}{automatic at Mechanic 1}{}{ I. Allows the character to attempt to pick locks and set or disarm traps.}


% Survival
\subsection{Survival}

\abil{Natural Shelter}{passive}{survival 1}{}{The character can build a shelter using natural objects in 4 hours. This shelter reduces the difficulty of exposure while sleeping by 5.}

\abil{Geographer I/II}{passive}{automatic at Geography 1/4}{}{ I. Allows the use of geography knowledge + detection as an exploration skill, provided you have a map. II. Allows for professional map-making.}

\abil{Tracking}{passive}{6 XP, INT 2}{ Survival 2 }{Learn to track.}

\abil{Odor Repertoire}{passive}{8 XP, INT 2}{Bear Nose, tracking}{Recognizes smells on a success if the scent is familiar. Increases food searching and tracking by +3 and allows tracking by smell, which may have a completely different DL to regular tracking.}

% Combat
\subsection{Cunning}

% \abil{Desperate Action I/II/III}{1 STA + action}{6 XP, INT 2/4/6}{}{Can perform a block, evasion, or reflex even without sufficient AP, but must spend 1 STA + the normal cost of the action to do so. Cannot end up with less than -2/-3/-4 AP. }

% \abil{Assassination }{passive}{6 XP, INT 3}{}{Increases the bonus to hit vulnerable attacks against immobile opponents or from stealth by +2.}

% \abil{Ambusher }{passive}{6 XP, INT 3}{}{When making stealthy melee attacks with a small weapon, if the first attack is a critical hit, you can attack again before the opponent makes their detection check or has any reaction. }

% \abil{ Strategist }{passive}{8 XP, INT 3/6}{}{ Increases the amount of AP from the utility burst by +1/+2. }

% \abil{Lightning Reaction}{passive}{8 XP, INT 3/6}{}{ Increases the amount of AP from the reaction burst by +1/+2. }

% \abil{ Strategist }{passive}{8 XP, INT 2/4/6}{}{ Can conserve 1/2/3 AP that remain from the previous turn for the next turn, exceeding the character's maximum AP. }

% \abil{Danger Intuition}{passive}{6 XP, INT 6}{}{The character is immune to starting combat surprised. }

\abil{Tactical Positioning}{3 AP}{6 XP, INT 5}{}{ Can spend up to 3 AP on movement, cunning, or prestidigitation actions at the start of combat, before any character takes their turn. This ability does not work if the character is surprised. } 

\abil{ Prepared }{passive}{6 XP, INT 3}{}{ The first time a prepared action is activated, it costs 1 AP less. }

% \abil{Last Breath}{5AP+3STA}{6XP, INT 4}{}{ The character can gain 3 AP, but spends 2 STA to do so and, as soon as their turn ends, loses 5 AP, potentially going negative.}

% Stealth

% \abil{Stealthy I/II/III}{passive}{4 XP, Subterfuge 2/4/6}{ PER 12/15/18 }{+2/4/6 Stealth.}

% \abil{Skilled Hands I/II/III}{passive}{4 XP, Subterfuge 2/4/6}{ PER 12/15/18 }{Gives +2/4/6 to prestidigitation. II. prestidigitation checks can be made safely.}

\abil{Guerrilla Tactics}{passive}{6 XP, INT 4}{}{When making ranged attacks stealthily, reduces the penalties for stealth to 3. If the character is downed, they do not receive the penalty for shooting.}

% \abil{Echolocation}{passive}{6 XP, INT 4}{Bat Ears}{Allows the character to see with their ears. Vision counts as dazzled, but it is omnidirectional. }

% \abil{Superior Attention I/II/III}{passive}{6 XP, Subterfuge 2/4/6}{Subterfuge 1/3/5}{I. Increases detection by +2/4/6.}

% \abil{Traveler I/II/III}{passive}{6 XP, Subterfuge 2/4/6}{PER 12/16/18}{Increases Exploration by +2/4/6.}

% \abil{Master Rogue I/II}{passive}{12 XP, Subterfuge 5/6}{}{ +1/2 to detection, stealth, and prestidigitation.}

% \abil{Supreme Intuition I/II}{passive}{12 XP, Subterfuge 5/6}{}{ +1/2 to cunning and exploration. }

% Smell

\section{Witchcraft}

\subsection{ General }

\abil{Magic Theory}{passive}{10 XP, INT 1}{Literacy}{ Grants the wizard the ability to maintain a grimoire, which allows for precise spell instructions to be written. Enables learning spells from grimoires. Can attempt to cast any spell written in a grimoire in a risky manner if out of combat. The difficulty to cast the spell is increased by 3 for each point of knowledge below the requirement. }

\abil{Battle Mage I/II/III}{passive}{6 XP, INT 2/4/6}{Melee Combat 1/3/5}{I. Allows the wizard to move while concentrating, repurposing the same action points used for the spell. II. When interrupted, instead of losing the spell, redo the spellcasting test with a penalty of -3 and choose the worst result between the original and this one. III. Allows the use of active defense while concentrating. }

% \abil{Mage I/II/III}{passive}{6 XP, INT 2/3/4}{INT 12/15/18, Magic Theory}{Increases witchcraft by +1/2/3. }

% \abil{Magical Practices}{passive}{4 XP, INT 3}{}{Quickly casting spells increases the casting difficulty by 2, not 3. Additionally, it is possible to use a spell with a cost of action points less than or equal to 2, after being accelerated, as a reaction when an opponent attempts to enter melee range.}

% \abil{Laws of Telekinesis}{passive}{4 XP, INT 3}{}{Extending the range of a spell only increases the casting difficulty by 2. }

% \abil{Magical Stability}{passive}{4 XP, INT 3}{}{Extending the duration of a spell only increases the casting difficulty by 2.}

% \abil{Efficiency Optimization I}{passive}{4 XP, INT 4}{}{Careful casting reduces the stamina cost of spells by 1.}

% \abil{Control Theory I/II}{passive}{6 XP, INT 3/5}{}{Overloading spells only increases difficulty by 4/3.}

\abil{ Favorite Spell }{passive}{6 XP, INT X}{}{ Choose a spell to reduce its difficulty by 2. The required talent is the same as that used to learn the spell. The spell can be performed safely. }

% \abil{Subconscious INT I/II/III}{passive}{6 XP, INT 2/4/6}{}{Casting hidden spells increases the difficulty of the sensing test by +2/4/6.}

% \abil{Magical Trance}{modal}{6 XP, Witchcraft 3}{}{ The character becomes unable to perform any action other than sensing or witchcraft. In return, the character's maximum action points are equal to INT. This ability must be activated at the beginning of a turn and cannot be turned off until the beginning of the next. }


\subsection{Animancy}

%Animancy
\inna{Synesthesia I/II/III}{0 AP + 1 STA per turn or 2 exhaustion per turn}{6 XP, Animancy 1/3/5}{}{ Animancy+Sorcery meter radius}{instantaneous}{ I. Maintains an energy perimeter that amplifies perception. This character can see invisible things as if they were visible within a sphere around them. The volume of perception is filled as if it were gas. Solid impermeable objects block the radius of synesthesia. II. Can maintain synesthesia without spending STA per turn and allows extending the range of enhanced senses up to 2x Sensing by spending 1 STA per turn. III. Doubles the normal and increased radius. }

\abil{Blind Combat}{passive}{6 XP, INT 3}{Synesthesia}{ Uses synesthesia as an active sense. Does not receive any distance penalties and does not receive any penalty for using a closed helmet while synesthesia is active against characters within its range.}

\abil{Martial Spirit}{2 STA}{6 XP INT 3}{Spiritual Attack}{}{ Makes the spiritual attack a favorite spell. Learns to cast spiritual attack along with melee attacks that cost at least 3 AP as a touch spell without increasing the cost of AP. Also learns to do the same with ranged attacks. }

\abil{Magical Interpretation}{passive}{6 XP, INT 3}{Synesthesia}{ Learns to identify spells cast by anyone within the area of synesthesia before they are cast, as long as they know the spell. }

\abil{Advanced Possession}{passive}{8 XP, INT 5}{ Possession}{ Makes possession a favorite spell. During a possession, can regenerate 1 STA per hour. }

\abil{Complete Possession}{passive}{x XP, INT 6}{ Possession}{ Can permanently possess a body. The XP cost of the ability is for possessing a body, where x represents the difficulty of possessing the body.}

\abil{Avatar}{6 AP}{10 XP, INT 4}{Synesthesia}{Allows a spirit to possess the character's body. Can utilize all skill and proficiency bonuses of the spirit. Every time the spirit and the player disagree on an action or the player wants to expel the spirit from the body, a will test must be made. Whoever succeeds or crits will do things their way.}


% end -----------------------------------------------------------------------

\subsection{Combustion}
%Combustão
\abil{ Micro-combustion intracellulare }{passive}{6 XP, Combustion 2}{ intoxicated}{ Gains +5 to resist cold damage while intoxicated. }

\abil{Grenadier}{passive}{8 XP, Combustion 5}{}{Learns to cast spells into grenades while throwing them, using the higher cost of AP between the throw and the spell. }

\subsection{Electromagnetism}

\abil{ Magnetic Shield }{ 1 AP + 1 STA }{6 XP, Electromagnetism 4}{ magnetic armor}{ Attracts metallic objects to the most resistant point of the armor. The opponent must make a strength test against the DL of sorcery to detach a metallic object from the armor. }

\abil{Electric Skin}{passive}{6 XP, Electromagnetism 3}{Taser}{ Turns Taser into a favorite spell. Taser can be accelerated to cost 0 AP and can be used with the hands, allowing the electric object to be stored.}

%Entropy
\subsection{Entropy}

\abil{ Walk on Water}{2 AP + 1 STA per turn}{6 XP, Entropy 2}{ Entropy 3}{ Can increase the surface tension of the water beneath your feet to walk on it. }

\abil{ Leidenfrost Barrier }{ 2 STA }{6 XP, Entropy 4}{ Combustion 4, Medicine 1}{ During rest, protects the face and respiratory system from environmental burn effects for one turn, preventing 5 damage and allowing breathing and keeping eyes open if no damage is received. }

\subsection{Radiation}

\abil{Sensory Practice}{passive}{4 XP, sorcery 3}{Radar}{ Radar becomes a favorite spell and can be kept active constantly for 2 Exhaustion per hour.}

\subsection{Biomancy}
%Biomancy

\abil{Blood Magic}{4 IL}{6 XP, Biomancy 5}{}{Receive 4 IL in exchange for health/2 STA.}


\abil{Cell Reconstruction I/II}{4 AP + 1 STA}{8 XP, Biomancy 2/4}{}{I. Spends 1 STA to accelerate the effect of a healing potion by an amount equal to health. II. Reduces the cost by 1 AP. }

% \abil{Powerful Voice}{2 AP, 2 STA}{6 XP, Sorcery 2}{}{Speaks very loudly. }

% \abil{Toxic Sweat}{passive}{6 XP, Sorcery 1}{ intoxicated, Entropy I}{ Gains +3 to resist exposure to insects while intoxicated. Receives +3 on any defense against swarm-type opponents, unless they have resistance to alcohol. }

% \abil{Mutagenic Agents}{passive}{6 XP, Sorcery 4}{Medicine 5}{Allows learning metabolic mutation skills with sorcery, but only on oneself. }

% \abil{Animal Grafts}{passive}{6 XP, Biomancy 4}{}{Allows performing grafts on plants and animals, enabling the learning of chimera mutation skills with sorcery.}


% Telepathy
\subsection{Telepathy}

\abil{Universal Communication}{passive}{6 XP, Telepathy 2}{}{Allows asking yes or no questions to any creature with 3 INT or more. The questions must be simple enough for the creature to understand.}

\abil{Central Intelligence}{2 AP per round}{8 XP, Telepathy 5}{Telepathic Link}{All characters connected by a telepathic link can choose to use the caster's cunning value instead of their own.}

\abil{Mind Reading}{passive}{6 XP, Telepathy 3}{Telepathic Link}{Allows reading the surface flow of thoughts in a telepathically connected mind, enabling the perception of images and hearing words they think and memories they recall. Gains +5 cunning against anyone under the telepathic link.}

\abil{Telepathic Suggestion}{passive}{6 XP, Telepathy 3}{}{ Makes suggestion a favorite spell and allows making suggestions telepathically to any detectable target, without the need to speak or establish a telepathic link. }

% \subsection{Transfiguration}

% Allows performing mutations on oneself and others.

% \abil{Mutagenic Agents}{passive}{6 XP, Sorcery 4}{Medicine 5}{Allows learning metabolic mutation skills with sorcery, but only on oneself. }

% \abil{Animal Grafts}{passive}{6 XP, Sorcery 4}{Medicine 5}{Allows performing grafts on plants and animals, enabling the learning of chimera mutation skills with sorcery.}

\section{CON}

% \textbf{Mutation Level:} Skills that have Mutation +x in their cost increase the character's Mutation level.

\subsection{Physical Transfiguration}

These are mutations that involve resizing the body. They can be learned instinctively through athletics or intellectually trained with transfiguration magic. It is possible to train this skill on another character, like plastic surgery, but the wizard must have the talent, and both must spend the necessary time to gain XP.

\abil {Grow! } {passive}{12 XP, CON 4}{STR > 15 }{ A humanoid character increases one size class and loses 3 STR and 1 AGI.}

\abil {Shrink! } {passive}{12 XP, CON 4}{ STR < 9}{ A humanoid character decreases one size class and gains 3 STR and 1 AGI.}

\abil{Large Hands }{passive}{4 XP, CON 4 }{racial }{ Allows the use of melee weapons from a larger size category with a -2 penalty to hit, an increased AP cost of +1, and with STR-3 in the contribution to weapon damage. The hands also count as a weapon of larger size. }

\abil{Keen Eyes I/II/III}{passive}{4 XP, CON 1/3/5}{}{Improves vision by +1/2/3. III. Reduces vision penalty from distance to -1 per band.}

\abil{Acute Ears I/II/III}{passive}{4 XP, CON 1/3/5}{}{Improves hearing by +1/2/3.}

\abil{Sensitive Nose I/II}{passive}{4 XP, CON 1/3}{}{Improves smell by +1/2/3.}

% \abil{Agile Build}{passive}{16 karma }{racial }{ Running speed increases by 1m per AP and gains +1m in height and long jump distance. All bonuses are revoked when armor penalty is higher than 1.}

% \abil{Strong Build}{passive}{16 karma}{racial}{Increases RES by 1, decreases armor penalties and maximum AP by 1. However, standing up costs 3 AP, and running and jumping have speed reduced by 1m.}

\abil{Beautiful}{passive}{4 XP, CON 4}{chosen at character creation or ritual}{ This ability can be used to gain +1 on persuasion attemtps and +3 on seduction attempts when allowed. This ability is used when allowed by the GM, meaning that beauty may have no effect against animals or in other situations where the character's appearance is not relevant. }

% \abil{Ugly}{passive}{8 karma}{chosen at character creation or ritual}{ Gains -2 on seduction checks and +2 on intimidation in social actions. }

\abil{Obese}{passive}{8 XP CON 1}{}{ Receives -1 load penalty to AGI. When Hunger reaches 45, the character loses this ability and their hunger resets to 15. }

\abil{ Visual Acuity }{passive}{8 XP, CON 5}{}{I. Reduces the distance penalty for cunning by half. }

\subsection{Implants}

The power of the abilities depends on the quality of the implant. Higher quality implants cost more.

\abil{Ignite Implant}{passive}{50 PO}{}{Allows the use of combustion abilities. }

\abil{Electricity Implant}{passive}{50 PO}{}{Allows the use of electricity abilities. }


\subsection{Spiritual Mutation}

These mutations are spiritual and require knowledge of Animancy or divine assistance to be acquired.

\abil{Elf}{passive}{ 10 XP, SPI 6}{Devotion 5 }{ Gains the ability to meditate for 4 hours a day to gain +3 Devotion for the day. The bonus also applies if in a sacred place. The character must meditate every day; otherwise, they will become ill within 24 hours. The entity worshiped by the cleric has +10 on the DL of compulsions against the elf's will. }


% end -----------------------------------------------------------------------

\subsection{Metabolic Mutations}

These are mutations that occur at the cellular level. They can be learned instinctively through athletics or intellectually trained with transfiguration magic. 

% \abil { Red Blood Cell Saturation I/II } {passive}{ 8 XP, CON 2/4} {} { Increases STA regeneration from resting by 1/2. }

\abil{Metabolic Control I/II }{passive}{10 XP, CON 2/4}{}{ I. Learns to trade regenerating 1 exhaustion for 1 healing during rests. II. Learns to perform multiple rests per day.}

\abil{Regenerative Reservoir }{passive}{8 XP, CON 4 }{Cell Reconstruction}{ The character has an internal reservoir that functions like a potion. The reservoir fills with rest while they are uninjured and can hold up to 6 heals.  }

\abil{Resurrection}{ritual}{10 XP CON 6}{ Metabolic Control II, Health 5}{ The body performs rests after the character dies. If the character remains without deadly injuries during this time, they come back to life. Resurrection does not heal amputations. }

\abil{Glutton}{passive}{8 XP CON 5}{ racial or training}{The character can eat 2x the normal amount per day.}

\abil{Tireless}{passive}{8 XP CON 5}{racial or training}{ Can rest in only 1 exploration turn.}

\abil{Hibernation}{passive}{8 XP CON 6}{racial or training}{ Can rest for long periods. The character only receives 1 hunger and 1 thirst every 2 days of sleep. }

\abil{Amplified Pheromones}{passive}{6 XP, CON 3}{}{ The character gains awareness and control over their pheromones, increasing the difficulty of any will or social cunning test by +2 for anyone within 6m.}

\subsection{Chimerism}

These are abilities that mix characteristics of other animals with those of humans. These mutations only occur when trained with transfiguration magic. It is possible to train this ability in another character, like a plastic surgery, but the wizard must have the talent and both must spend the necessary time to gain XP.

\abil{Fur / Scales}{passive}{CON 3}{racial or ritual}{Choose Fur or Scales. Fur: Becomes Furry, granting +3 on tests to resist exposure to cold and increases natural INS by 2xDM. Scales: Makes the skin resemble that of a crocodile. Increases natural TGH by +2xDM.}

\abil{Retractable Claws}{passive}{6 XP CON 3}{racial or ritual}{Gives an unarmed attack with a base damage of 6/1x and the properties of grappling and heavy I. Also improves climbing by +2 if not wearing gloves.}

\abil{Gills}{passive}{10 XP CON 6 }{racial or ritual}{Can rest with half the STA regeneration underwater and does not lose STA due to suffocation.}

\abil{Tail I/II}{passive}{5 XP CON 3/4 }{racial or ritual}{+1/2 for balance and climbing. II. Can hold things and hang with the tail.}

\abil{Bear Nose}{passive}{10 XP CON 5}{racial or ritual, sensitive nose II}{ Modifies the nose to become very sensitive and can transform it into a snout. Removes the -5 penalty to smell.}

\abil{Infrared Vision}{passive}{10 XP CON 5}{racial or ritual}{Modifies the eyes to be able to see infrared light, allowing aiming at heat sources even through fog or in the dark.}

\abil{Owl eyes}{passive}{10 XP CON 5}{racial or ritual}{Modifies the eyes to be able to see well at night, but not absolute darkness. }

\abil{Bat Ears}{passive}{10 XP CON 5}{racial or ritual}{Modifies the ears to pick up a wide range of frequencies and become more sensitive. Turns ears into an active sense. This is not as good as vision as a dominant sense, so it leaves the character disoriented when used as the only active sense. }

\abil{Four Legs}{passive}{12 XP CON 5}{racial or ritual}{ Gains quadrupedal movement, gaining +3 on any balance test and its running speed is AGI/2. Cannot have anything in its hands while walking on four legs. }

\abil{Venom Glands}{passive}{10 XP CON 5}{racial}{Allows for performing abilities with poison.}

% \inna{Blinding Spray}{4AP + 4 SP}{forest 3}{melee, range of 4.5m, arc of 180 degrees}{3 turns}{Sprays an acidic cloud that burns the eyes of anyone. If they fail a Health test, they will be blinded for the duration. Alternatively, they can try to escape with a reflex to leave the area.}




%Charisma
% \abil{Collective Hysteria}{passive}{6 XP, Animancy 2}{ }{ Each character that is afraid or angry increases the DL of fear or anger spells by 1. }



% \abil{Performance Anxiety}{passive}{4 XP, Subterfuge 2}{}{Increases the DL of casting all spells by 5 against any enemy that is afraid.}

% \inna{Boogeyman}{4 STA per turn}{4 XP, Animism 4}{ }{10m}{instantaneous}{If an enemy is afraid, they will have to defend against the overwhelming terror spell when performing any offensive action against the caster.}

% Lust

% \abil{White Knight Syndrome}{}{}{}

% \abil{ Stockholm Syndrome Attack }{ 1 AP }{ 4XP}{ }{ When the bard hits someone with a melee attack, they can forgo the damage to try to charm the opponent with +2 on a hit and +6 on a critical.}

% Confusion

% \abil{Gaslighting}{}{}{}


\end{multicols}

