\chapter{Combat}

\begin{multicols}{2}

\section{How skills work}

Skills are things a character can do and improve upon. To measure how good someone is, a skill value is calculated.

\textbf{Skill test:} A skill test consists of a dice roll to determine the degree of success in performing an action. Tests are made by rolling a 10-sided die and adding the result to the skill value. This sum is compared to a difficulty level to determine the degree of success of the action.

\textbf{Difficulty level (DL):} A level that defines how hard it is to overcome something. It is the number against which skill tests are made to determine the degree of success.

\textbf{Degrees of success:} Critical, hit, graze, and failure are the degrees of success. Critical is the best success, and failure is the worst failure. Regardless of the type of skill test, a score that exceeds a DL or defense by 10 or more is a critical success, 5 or more is a hit, 0 or more is a graze, and anything less than that is a failure.

\textbf{Target number (TN):} Works like the DL, but the action has a hit or miss. In this case, if the test value is greater than or equal to the TN, the action is a success; otherwise, it is a failure.

\textbf{Exploding die:} In skill tests, if the result of the d10 is 10, roll 1d6 and add that to the score, continuing to add while the result is 6. If the result of the d10 is 1, do not add anything and then roll 1d6 and subtract the result instead of adding. Continue subtracting while rolling 6.

\textbf{Safe and risky tests:} It is possible to perform some tests safely or riskily. In the safe form, roll 2 d10 and take the result closest to 5. Similarly, it is possible to perform tests in a risky manner. In this case, roll 2 d10 and take the result farthest from 5. This mechanic can only be used when indicated in the skill or ability or when the Master allows it.

% \textbf{Special Failure:} If the total skill value is less than 5 due to a natural 1, something especially bad may happen, such as breaking a weapon, dropping all arrows from the quiver, or falling. This is an optional rule, and its application is entirely at the Master's discretion.

\textbf{Note on difficulties:} When designing DLs, to put things in perspective, consider that a DL of -5 is trivial, 0 is average, 5 is difficult, and 10 is very difficult.

\section{Turns and Actions}

\textbf{Start:} Combat occurs when a group decides to fight against another. A combat is played in rounds, which represent 3 seconds of game time each.

\textbf{Preparation:} At the start of combat, the miniatures are positioned in their locations, and each character starts with 6 AP by default. They also receive a turn token, which they can spend to take their turn.

\textbf{Surprise:} A combat can start with characters being surprised. In this case, all characters who are surprised begin the first round with 0 AP.

\textbf{Turn order:} It is defined by a "first to ask, first to play" rule. If several players want to play at the same time, all interested make a cunning test, and the one who rolls the highest goes first.

\textbf{Ending a turn:} Once all desired actions have been performed or prepared, a character can end their turn. They can save as many AP as they want.

\textbf{Action surge:} All characters are entitled to one action surge per round. There are 3 basic types of action surges: movement, combat, and reaction.

\begin{itemize}
    \item Movement: 3 STA to gain 6 AP that must be used for movement immediately and allow running until the end of the turn.
    \item Combat: 3 STA to gain 6 AP that must be used immediately.
    \item Reaction: 3 STA to gain 6 AP, but can only use reactions and rest until the end of the round. It is possible to rest and perform the reaction surge simultaneously, as long as AP and STA remain greater than or equal to 0 at the end of the action.
\end{itemize}

\textbf{End of the round:} A round ends when no one has AP left, nor action surge. When the round ends, all characters regain their AP. If anyone wants to end their turn without spending all AP, any AP that was not spent can be combined with that of the next round to rest, recovering STA.

\textbf{Environmental and ongoing effects:} Any effect that occurs periodically per turn or is environmental has its effect applied at the beginning of the round.

\textbf{Negative AP:} AP can become a negative number, but never as a result of performing an action or reaction. When the character recovers their AP, they must deduct the negative value from their maximum.

\textbf{Suffocation:} If the character cannot breathe during the turn of a round, they lose 1 STA and cannot rest.

% \textbf{Delay the turn:} The player can spend 1 AP during their turn to delay it. This allows the character to regain their token and play again within the same round. AP regeneration only occurs once per round.

\textbf{Reactions:} These are actions that can be performed on another character's turn but must be triggered by something. An example is trying to evade an attack, which is triggered when being attacked. A reaction takes effect before the action that caused it.

\textbf{Preparing a reaction:} Any action can be prepared on a player's turn, but doing so costs 1 AP. This allows the player to create any trigger they can imagine for a reaction. They must have enough AP to perform the reaction when the trigger occurs. A prepared action can consist of a movement and a use of skill and can be repeated multiple times as long as the trigger happens again.

\textbf{Compulsions:} These are forced actions. They must occur immediately if there are enough AP. If not, they force action surges. If even that is not possible, the remaining AP are lost, and the compulsion ends. If the compulsion occurs during another character's action surge, it happens when the AP of the surge ends.

% \textbf{Severe compulsions:} Whenever a compulsion forces a character to perform an act that they perceive will cause physical harm to themselves or someone they care about, the compulsion is severe. In this case, the character has +5 to resist the compulsion.

% \textbf{Compulsions:} These are forced actions. When the compelled character has enough AP, the action is performed immediately; otherwise, it occurs at the start of their next turn, right after regenerating AP and processing effects per turn. The turn of the compelled character must start as soon as possible. If the compulsion requires a reaction, such as defending someone, it is not possible to end the turn with fewer AP than the compulsion requires. If it is not possible to fulfill the compulsion within a turn, lose the compelled AP to get rid of it. The most recent compulsion overrides the previous one if there are 2 compulsions stored for the next turn.

\section{Movement}

Movement actions spend AP in exchange for changing the character's position on the grid. A character must always end their turn in the center of a square or hexagon, which means that moving only a fraction of a space is impossible. It is possible to pass through a space occupied by another character as long as they do not resist the passage. Multiple characters can occupy the same space, depending on their sizes, but they may suffer from the grappled condition.

Types of movement:
\begin{itemize}
\item \textbf{basic movement:} Simple walking movement.
\item \textbf{careful movement:} Slow movement used to improve stealth and balance.
\item \textbf{running:} Can only be initiated during a movement surge, but can continue running after accelerating as long as no turns of 90 degrees or more are made per running block. The character can run a partial amount smaller than its running speed, but AP cost is the same.
\item \textbf{getting up:} Removes the fallen condition.
\item \textbf{jumping:} A jump has a vertical distance equal to half the horizontal distance. If performed during a run, increase the horizontal distance of a long jump to AGI/3. Movement cannot be voluntarily interrupted in the middle of a jump. If performed backward, the horizontal distance is halved.
\item \textbf{crawling:} is the only usable movement speed while \textbf{lying down}.
\item \textbf{swimming:} is the only usable movement speed while swimming. It also applies the \textbf{lying down} condition.

\item \textbf{trample:} The character attempts to pass through an occupied space. If the opponent does not move out of the way, they must perform a push action to stop the movement.
\item \textbf{mounted movement:} A character needs minimum training to be able to use a mount. During mounted combat, the character's movement speed matches that of the mount, but they must spend an extra 1 AP to give a movement command, such as run and stop (not changing direction if the change is less than or equal to 60 degrees). Poorly trained mounts respond more slowly. Mounting and dismounting costs AP equal to standing up.
\end{itemize}

\begin{center}
\begin{tabular}{| c c c|}
\hline && \\
Movement & Cost & Speed  \\
\hline && \\
Careful & AP & 0.5 m \\
Basic & 1 AP & 1 m \\
Run & 2AP & AGI/3m \\
Jump & 1AP+1STA & AGI/4m \\
Crawl & AP & 0.33 m \\
Swim & AP & 0.33 m \\
Swim Fast & AP & 0.5 m \\
Stand up & 5 -AGI/5 AP & 0 \\

\hline
\end{tabular}
\end{center}

\skill{Swimming}{AGI/2 -5}{ Represents the character's swimming ability. Make a swimming check. On a critical or success, they can swim quickly for 8 AP per STA. On a success, they spend 1 STA for every 4m of fast swimming. On a near miss, the speed is normal. On a failure, the character spends 1 STA per AP of swimming, falls 1m, and their speed is equal to that of a near miss.

The swimming check should not be redone more than once per round unless the swimming conditions change.

Characters attempting to swim with one hand receive a -3 penalty on the check.

A swimming check in a deep pool is considered a DC of 8. Swimming at a beach with waves is 10. Swimming in a swift river or choppy sea would be 13.
}

\skill{Climbing}{AGI/2 -5}{ Make a climbing skill check to overcome a DC and climb a surface. Climbing checks can be made safely or recklessly.

The climbing speed is equal to the crawling speed but costs 1 AP + 1 STA. On a critical, it does not cost STA. On a success, they climb normally. On a near miss, decide to go twice as slow or give up. On a failure, the character simply falls from the surface.

A character who is climbing is considered lying down. An attack that interrupts the climber will require them to redo their climbing check with a -3 penalty. In the case of multiple climbing checks in a single turn, always take the worst result. Climbing with one hand incurs a -3 penalty.

To determine the difficulty of climbing, consider whether the surface is moving, the quality of the grips, and the distance to be climbed. An easy climb, with minimal chances of falling, is considered a DC of 5. A DC of 15 is reasonable for out-of-combat situations but quite challenging in a difficult scenario. DCs greater than 20 will require equipped and specialized characters and are unlikely to be accomplished during combat. As a general rule, consider that for every meter, the difficulty increases by 1.}

\skill{Balance}{AGI/2 -5}{ Skill to perform acrobatics, avoid falling, and mitigate damage from falling. Can be used actively with a skill check or passively when something tries to knock you down.

\textbf{Difficult terrain} Every time the character moves through difficult terrain, they must make a balance check. The DC is based on how slippery, unstable, long, and narrow the path is. The character decides the type of movement they wish to use, then makes the skill check. On a critical, all movement is inconsequential. On a success, moving at a normal speed is safe, but not jumping or running. On a near miss, moving at a careful speed is safe. On a failure, only crawling is allowed. If the character enters or moves in difficult terrain using a type of movement they did not manage well enough, they fall.

When falling, the DC to fall properly is arbitrary, but is generally higher for low falls from bad positions, average for very high falls, and lower for small falls from a controlled position. DCs should be around 20 at most and 5 at least.

\textbf{Fall damage} depends on the speed of the fall and the landing spot. The standard landing spot is a flat, hard surface. The speed of the fall can be considered instantaneous for falls below 2m, in which case it causes 10xDM of bludgeoning damage. 5m per second for the first 10m, causing 30xDM of bludgeoning damage. Then it increases to 15m per second up to 30m, causing 100xDM of damage. Anything above that causes 200xDM of damage. A critical on the balance check reduces the damage by 75\%, a success by 50\%, a graze by 25\%, and a failure takes the full damage.
}  

\section{Morale}

Morale is the measure of the characters' stress tolerance. Stressful situations, such as combat and exploration in terrifying places, trigger morale tests. The test is conducted using will proficiency.

\textbf{Moral test:} At the beginning of each round, all characters make a will test against the terror of the moment. A graze leaves the character frightened, and a failure activates the character's limit stress action.

DL = 0 + aggravating factors

Usual aggravating factors: \\ 

\begin{itemize}
    \item Sensory deprivation +5
    \item Seriously injured: +5
\end{itemize}

\textbf{Intimidation:} A character can make an intimidation attempt as part of an action in combat. This increases the value of the DL of the morale test by the character's charisma. The intimidating action is made as the player describes how they try to demoralize enemies. If they must do extra things, like gestures or talking, the action may cost AP to be performed. The amount is to be decided as the GM sees fit. Some abilities and creatures may automatically increase the DL of the morale test for one reason or another. It may be that the creature is being very aggressive, delivers a very painful, quick or spectacular death experience to its victims or that it is very unpredictable and that is what raises stress.

\textbf{Taunt:} In combat, a taunt is an attempt to make the target infuriated. It costs 2 PA to make and can be done by saying something, making faces, opening your guard deliberately or any other action that can be considered provocative. The target must make a will test against the attacker's persuasion. On a graze, the target is compelled to attack the taunter on their next action. On a crit, they are also enraged.


% Every time the character activates their limit stress action, they lose 2 XP from one of their convictions, at the character's choice.

\section{Damage and Injuries}

\subsection{Injuries and Death}

\textbf{Injury level (IL):} Any time a character receives damage, they will increase their injury level. They will receive injury penalties for every multiple of their \textbf{injury tolerance (IT)}. The base value for injury tolerance is 5.

\textbf{Injury Capacity:} is the maximum amount of injuries a character can take without collapsing and going unconscious. Its value is 8x IT.

\textbf{Death:} occurs when the character's injuries exceed 10x IT.

\textbf{Healing Injuries:} characters heal by resting well and using healing magic or potions. Any time the injury level is healed, the character must spend 1 hunger and 1 thirst unless otherwise stated by the specific healing method.

\textbf{Damage:} Any time a character receives damage, it must compare the amount received with the respective threshold for the damage type to determine the injuries caused. 

\textbf{Damage Thresholds:} 
\begin{itemize}
    \item Tier 1. Damage is higher than the armor's protection to specific type of damage. Increases IL by +1.
    \item Tier 2. Damage is higher than tier 1 + character's resistance to that kind of damage. Increases IL by +5.
    \item Tier 3. Damage is higher than tier 1 + 2x character's resistance to that kind of damage. Increases IL by +10 and has a 25\% chance of causing a wound.
    \item Tier 4. Damage is higher than tier 1 + 3x character's resistance to that kind of damage. Increases IL by +20 and is guaranteed to cause a wound.
    \item Tier 5. Damage is higher than tier 1 + 5x character's resistance to that kind of damage. Causes instant death or amputation.
\end{itemize}


\textbf{Types of damage:} are: physical, burn, electric, radiant, and corrosive. Objects and creatures may have resistance or vulnerability to each of these types of damage. Different types of damage cause different kinds of wounds.

\textbf{Physical damage:} There are two components to consider when receiving injuries from physical attacks: Impact (IMP) and penetration(PEN). Attacks and effects that cause injuries will have some amount of each of these components. 

\textbf{Penetration:} is the component that represents the ability to cut the target. Sharper and harder weapons have greater penetration. Its respective defense is TGH. If the attack causes an injury from penetration, it ignores the armor's insulation and increases the DL of any poisons by 3 for serious injuries and by 5 for deadly ones.

\textbf{Impact:} is the component that represents the total force applied with the blow. It s respective defense is RES.

\textbf{Interruption:} is an effect that causes the victim to lose their action and associated costs when it occurs during an action or causes a minimum loss of 2 AP. Interruption occurs whenever a character receives a tier 2 or higher damage from impact or electric damage. Multiple simultaneous interruptions cause the AP loss of all interruptions combined or the AP loss from the lost action, whichever is greater.

\textbf{Burn, radiant, and corrosive damage:} Their respective defense is INS. These kinds of damage increase the IL by +5 for every damage tier. For example, tier 2 gives +5, tier 2 + INS gives +10, tier 2 + 2x INS gives +15 and so on. There is no sudden death.

\textbf{Electric damage:} Also defended by INS, but receiving a tier 3 damage from electric damage causes an interruption of 5 AP, and tier 4 leaves the target unconscious.

\textbf{Wounds:} can be caused by receiving tier 3 or 4 damage. Their healing counter is separate from the main ijury counter. 

\begin{itemize}
    \item Broken arm: makes the arm useless. Takes 10 cures to heal.
    \item Broken leg: makes the leg lame and useless. Takes 15 cures to heal.
    \item Hemorrhage: a special wound that requires 2 cures to stop bleeding. For every STA spent, the character receives +1 to their IL per hemorrhage cure untreated. It is caused by damage to the chest or head. Hemorrhages do not require the expense of hunger or thirst to be healed.
    \item unconscious: causes unconsciousness.
\end{itemize}

\textbf{Armor coverage and Precision:} Attacks have the possibility of hitting an unprotected part of the body, having its PEN ignore armor toughness. To determine whether that happens, add the weapon's precision bonus and subtract the armor coverage from the attack test result. The new result is used to determine the degree of success of the attack. The decision to use this feature can be done after rolling the dice, but it is irrevocable. 

\textbf{Localized damage and vulnerability:} Each type of creature may have points that must be specifically aimed at when using the attack, called vulnerabilities. When someone declares that they are aiming at that specific point, the attack is modified according to the description of the vulnerability. The standard place to hit if no location is specified is the chest.

For humanoids, there are 3: hand, leg, and head.

\textbf{Hand:} Has no armor unless it has gauntlets. A wound renders the hand useless. Ranged attacks have -5 to hit. Maximum IL damage taken from this is 5. A tier 5 damage amputates.
\textbf{Leg:} A wound leaves the character lame. Maximum IL damage from this is 10. A tier 5 damage amputates.
\textbf{Head:} receives -5 to hit, but the injury tier is increased by 1, except if the increase is from 1 to 2. Attacks that cause interruption and are tier 3 or 4 after increased leave the character unconscious.

Some creatures may have their vulnerable point out of reach of others, requiring long weapons, jumps, and proper positioning to reach. These special cases can be added where the Master deems it makes sense.

\subsection{Afflictions}

Afflictions are the negative conditions that can affect creatures. 

There are 5 categories of affliction: Mobility, Injury, Vision, Mental, and Health. 

\textbf{Injury:} penalties affect AGI, STR and half the value to Will. The penalties affect STR, AGI and all of its derived values, except for the damage thresholds and movement speeds.

\textbf{Mental:} Penalties affect Exploration, Cunning, Sorceries, Will, Insight, and Knowledge.

The list of afflictions is:

\begin{itemize}
\item \textbf{Lame:} Cannot run, jump, or perform basic movement.
\item \textbf{Prone:} Can only crawl and stand up. Becomes vulnerable to crushing attacks.
\item \textbf{Grappled:} Movement requires pushing, and attacks can only be made with short-range or small weapons.
\item \textbf{Immobile:} Cannot move and cannot use any combat or movement skills other than escape strength or grapple. The SD is reduced to -5 + SMx1. 
\item \textbf{Dazzled:} Limits the radius of vision to 5m.
\item \textbf{Blind:} Cannot use vision with Cunning.
\item \textbf{Deaf:} Cannot use hearing with Cunning. 
% Requires a balance check for basic movement on any terrain.

\item \textbf{Unconscious:} The character cannot do anything other than Will and Health checks. They can also make detection and sensing checks, but receive -10 on these checks. These checks can be made to wake up. The DL and the allowed check depend on the situation. 

\item \textbf{Afraid:} Receives -1 Mental and cannot make a fight surge. The fear ends when the character succeeds on their Will check or becomes angry. 
\item \textbf{Enraged:} Receives -1 Mental and cannot make a reaction surge. The anger ends when the character succeeds on their Will check or becomes afraid. 
\item \textbf{Confused:} Receives -3 Mental. Becomes vulnerable to domination attacks. 
% \item \textbf{Charmed:} Receives -3 Mental against their charmer. 
\item \textbf{Dominated:} Under the control of another character. Must obey commands for the duration but can make a Will check whenever the compulsion is severe. Outside of combat, can make a check once per hour.
\item \textbf{Distracted:} Receives -2 Cunning.

\item \textbf{Intoxicated I/II/III} -1/2/3 Mental.
\item \textbf{Tired/Exhausted} Penalty of -1 mental/-2 mental. 
\item \textbf{Weakened/Malnourished} -1/-2 Health. +5/10 to IL
\item \textbf{Thirsty/Dehydrated} -1/-2 Health, +5/10 to IL
\item \textbf{Sick} Counts as +10 to IL. -2 Health.
\end{itemize}

\skill{Health}{ CON }{ Represents how strong the character's body is against diseases, poisons, exposure to the elements, and how quickly they recover from injuries and exhaustion. 

The survival chapter describes how health works during rest.
}

\subsection{Melee Combat}

%---------------------ATTACK

\skill{Strike}{Melee}{ Striking consists of a melee weapon attack on a target using the strike skill test against the opponent's defense. The AP cost depends on the weapon used and the type of attack performed. Hits deal full damage, and crits deal 150\%. Grazes and misses have their effects determined by the type of defense used.

\textbf{Attack against equipment:} An attack can be made against a weapon, shield, or any object in order to break them. The shield deflection bonus is not added, and the impact is 150\%, 100\%, 50\%, and 0 in the case of critical, hit, glancing blow, and miss, respectively. The multiplied impact applies to the target object and the base damage to the weapon.

Other types of melee attacks will be available depending on the type of weapon used. The type of attack must be declared when used. Opportunity, wide, and heavy attacks can be used together. The others must be used individually.

\textbf{Heavy Attack:} It is possible to make a heavy attack with weapons that have the "heavy" property. Heavy I adds +0.5xSTR to the impact and increases the cost by +1AP+1STA. Heavy II adds +1xSTR to the impact, costs +2AP+1STA, and gives a -2 penalty to hit. Heavy III adds +1.5xSTR to the impact, costs +3AP+1STA, and gives a -3 penalty to hit.

\textbf{Sweeping Attack:} spends 1 additional AP on the attack and hits everything in a semicircle. Decide whether to attack from right to left or vice versa and attack the enemies in that order in a 180-degree arc. If the attack is blocked or redirected, the impact is reduced for all. If one character is behind another, only the closest is hit. Sweeping attacks do not hit fallen enemies. There is no size bonus or penalty to strike in sweeping attacks.

\textbf{Opportunity Attack:} A melee weapon attack that can be used as a reaction when an opponent attempts to perform a triggering action. It is possible to cancel the triggering action and reuse the AP spent to defend against an opportunity attack, but not with evasive jump. The attack requires the normal AP cost and has a +2 bonus to hit. Triggering actions include casting spells that cost 3 AP or more, using ranged weapons, using a standard action, standing up, and moving towards a melee weapon while within its attack range. The attack occurs before the effect of the triggering action, unless stated otherwise.

\textbf{Braced Attack:} A melee attack that attempts to increase strength by firmly bracing the weapon. It can be used as a reaction when a target is running towards the weapon or as an action in situations where the attacker is mounted. It costs an additional +2AP and +1STA on top of the normal attack cost and gives a -2 penalty to hit. The impact of this attack has a +STR bonus and knocks the opponent to the ground on hits if it is greater than 2x the target's natural RES. On a critical, it knocks down if impact > the target's natural RES.

\textbf{Knockdown:} A melee attack with a hook weapon that attempts to knock down the opponent. The attack costs an additional +1STA. If it is a critical, it knocks down the opponent with impact > natural RES. If it is a hit, it knocks down with impact > 2x natural RES. Gains +0.5xSTR if the opponent attempts to perform an evasive jump or is running. This can be used as a reaction against mounted targets that move away from the weapon within its melee range.
}

% \textbf{Disarm:} A melee attack that attempts to disarm the opponent. This attack does not deal damage, but if it is a critical, it disarms the opponent with impact > natural RES. If it is a hit, it disarms with impact > 2x natural RES. Gains +0.5xSTR if the opponent attempts to block by redirecting.

%-----------------------------------DEFEND

\skill{Defend}{Melee}{ This is the defense against the character's melee attacks. Whenever attacked, the character must spend AP to use the defend with some object or part of the body; otherwise, they use the SD. If the strike hits or crits, the defend does nothing. If it glances or misses, the result depends on the type of defense chosen. There are four types of defense: Evade, Evasive Jump, Redirect, and Block.

\textbf{Evade:} The defender spends 2AP to try to get out of the way of an attack. On a glancing blow, the attack has 50\% damage, and on misses, it has 0. If the character does not react to defend, they automatically use evade, but with the SD as the skill value. It is possible to receive -3 on the defend to evade with only 1 AP.

\textbf{Block:} Costs 2 AP, but only shields can be used to block precise attacks. On a glancing blow or miss, it reduces the PEN by the TGH and the impact by the RES of the blocking object. The remainder is transferred to the character.

\textbf{Redirect:} makes a defensive attack against a weapon by spending 3 AP. On a glancing blow, the attack impact is reduced by STR for one-handed weapons and by 2xSTR for two-handed weapons or shields. On a miss, the impact is further reduced by +1xSTR. The resulting impact hits the defender and is used to calculate the PEN. Redirect has a +3 bonus against long weapons if the characters are at close range.

\textbf{Evasive Jump:} is the most defensive option. It costs an extra 1 STA and allows the character to jump from the attack while blocking or evading an attack. This movement cannot be followed and gives +AGI/3 on the skill test. This can only be used if there is space to jump. The jump can be backwards or sideways and uses the movement speed of jumping backwards.

\textbf{Protect:} It is possible to defend an ally by staying within their melee range and the attack at the same time. The defender can redirect or block an attack directed at an ally.}

%----------------------- Grapple

\skill{Grapple}{Melee}{ Used to grab and subdue others. When a character is under the effect of grapple, it cannot move without dragging the other grappler. It also receives mobility penalties associated with the grappled condition. Additionally, the list of things a grabbed character can do is limited to the points below:

\textbf{Initiate the Grab:} attackers must make a strike test with a weapon that haS the grapple property against the opponent's defense or SD. On a hit or critical, the opponent is grappled, and the evasive jump movement is canceled. If the opponent does not attempt to defend the attack. If a creature's grapple skill is 10 points higher than its oponents, it does not suffer the grappled affliction.

\textbf{Attack and Defend:} Strikes between opponents involved in a grapple can be done with short range attacks.

\textbf{Grapple Maneuvers:} These are performed during a grapple by any of the participants in order to \textbf{escape, move, immobilize, disarm, or knock down} an opponent. These actions cost 3 AP +1STA, must be defended with the grapple skill, and require the defender to spend 2 AP+1 STA or suffer a -5 penalty. If the grapple attack has any damage, it deals that damage whenever a grab maneuver is used against the opponent.

\textbf{Drag and Push:} Grabbing someone enables the actions of dragging and pushing, which use strength as the skill.

\textbf{Escape:} Escapes from the grab on criticals and hits. Being interrupted allows for a reaction to escape without the possibility of active resistance.

\textbf{Disarm:} Removes something from the opponent's hands on a critical. It is possible to grab the opponent's weapon on a hit, preventing them from using it until they manage to escape.

\textbf{Knockdown:} The opponent falls to the ground on a critical. It is possible to throw oneself along to achieve a knockdown on a hit.

\textbf{Immobilize:} The opponent becomes immobilized on a critical. It is possible to stay immobilized yourself to achieve immobilization on a hit.

Multiple characters can assist each other in the grab. Any skill test made by several characters uses the highest skill value among them plus 4/3/2/1 for each additional character, up to a maximum of 5. All characters involved must spend the necessary AP. Characters of a smaller size category add a maximum of 2.
}

%----------------------SHOOT

\subsection{Ranged Combat}

\skill{Accuracy}{ Ranged Combat }{ Accuracy consists of a throw or shot directed at a target using the Accuracy skill test in ranged combat against the opponent's reflexes or their SD, should they choose not to react to the attack. The cost in AP depends on the weapon used and the type of attack performed. Hits deal full damage, and crits deal 150\%. Grazes and misses have their effects determined by the type of defense used.

\textbf{Distance Penalties:} If the target is 10m away or less, there is no penalty. Beyond that, a cumulative penalty of -2 to the attack is applied for every additional 10m of distance. From 50m onwards, the penalty is -2 for every additional 50m. The vertical range of a ranged weapon is half the horizontal range.

\textbf{Distance Measurement Recommendation:} Measuring distance is a time-consuming process, so it is more practical to estimate the distance intuitively and give the benefit of the doubt in favor of the shooter.

\textbf{Canceling Attack:} It is possible to evade, dodge, or block while shooting, but the shot is canceled, and the costs of AP and STA are the greater of the defensive action or the shot.

\textbf{Immobile Targets and Shooting Errors:} The DL to hit an immobile target occupying a space of 1m is equal to -5-2*SM. If a projectile misses the test, it hits some adjacent space or a distance even greater for every 3 points the test is below the DL.

\textbf{Moving Targets:} It is possible to shoot at moving targets as a reaction. The total amount of AP spent to shoot cannot exceed the amount of AP spent on movement. The AP count starts when the target becomes visible. The target's position at the moment of the shot is where it is at the end of the movement.

\textbf{Prepared Shots:} If a character prepares a ranged attack, the shot occurs when the target becomes visible.

\textbf{\large Types of Ranged Attacks:}

Other types of ranged attacks will be available depending on the type of weapon used. The type of attack must be declared when used, and only one can be used at a time unless stated otherwise.

\textbf{Throw:} Basic ranged attack with a throwable weapon. Throwable weapons gain range based on STR.

\textbf{Shoot:} Basic ranged attack with a shooting weapon. It cannot be blocked without a large or tower shield.

\textbf{Snipe:} Modifies the shot to cost an additional 2 AP to gain +2 in accuracy and shoot with a -1 distance penalty instead of -2 for the first 50m.

\textbf{Quick Shot:} Modifies the shot to cost 1 AP less to shoot, but incurs a -3 penalty in accuracy and -3 in distance instead of -2.
}

\textbf{\large Explosions:} This type of attack results from something like a bomb, which radiates from a central point and loses strength with distance. Explosions are defended against with a reflex test. If the explosion comes from a projectile, the DL of the explosion is equal to the shooting skill. If the explosion occurs before being perceived, no test can be made. If the bomb is perceived, players receive a certain amount of AP to move away before it occurs.

Explosions have different magnitudes of effects depending on where things are in relation to the center. The center is where the axis of the spell's radius is located. For a circle, it is in the middle. For a cone or line, it is in the first space from where the explosion radiates. Anyone caught at the central 30\% of the radius receives the critical effect, while in the 30\% to 70\% range they are hit, and in the outer 30\% they are grazed. Always round the middle area up and the others down. If a creature occupies multiple spaces, apply the strongest effect. In most cases, only the central space will be a crit and only the outer ring will be a graze. It changes once radius goes over 7m, but most explosions are not that big.

Effects of an explosion, such as pushing enemies, receive +5 against enemies in the critical area and -5 against those in the graze area. The damage from explosions is 200\% on a critical, 100\% on a hit, 50\% on a graze, and 0 on a miss.

\textbf{\large Sprays:} They are a type of explosion, with the same damage progression based on distance from the origin and the same DL definition mechanic. Sprays target all characters in a way that no two targets are more than 180 degrees apart in a circle centered at the origin of the spray. All targets make their reflex saves to escape, just like in explosions and move their characters according to their result. After the movement, the spray damages only one of the targets, at the attacker's choice.

%--------------------- REFLEX

\skill{Reflex}{ Ranged Combat }{ Used to react as quickly as possible to something. This can be used to avoid explosions, projectiles, and other effects that require quick reactions.

\textbf{Evasion:} When the character is targeted by projectiles, they can spend 2 AP to react. This AP cost is used to move and does not need to be spent again, but any STA cost must be paid. Upon receiving a critical or hit, take the attack normally and then use up to 2 AP to move if not interrupted. On a graze, the character can jump to try to gain cover before being hit and only takes half damage if they fail. On a miss, the character does not take the attack and can spend their movement surge immediately to escape or do the same as in the graze.

\textbf{Guard:} If using a shield, they can spend 2 AP to guard against ranged attacks coming from one direction as a reaction. All ranged attacks within a 180-degree semicircle in front of the character are made against the DL of reflexes + shield deflection without additional AP cost. The guard is canceled if the character performs any action other than careful movement.

\textbf{Avoiding an Explosion:} Spend 2 AP and make a reflex skill test against the DL of the explosion. On a critical, the character can run by spending one extra STA. On a hit, they spend an extra STA to jump in any direction before the explosion occurs. On a graze or miss, they can move 2 AP after the explosion.

\textbf{Explosion Block:} Can only be done with a shield. Make an active reflex + shield deflection test against the DL of the explosion. If successful or critical, they have cover against the explosion and reduce its damage by RES, TGH, and INS from the shield.

In all cases, if the character is interrupted, they cannot move after the interruption.
}

\subsection{Combat Positioning and Object Interaction}

\textbf{Flee:} It is possible to activate the movement surge as a reaction for a character to prevent them from entering melee range or getting any closer. The action must be performed immediately before the opponent enters range. Only the AP gained from the movement surge can be used with this reaction. 

\textbf{Follow:} A character can, as a reaction to any movement, follow another character who is within melee range. The action must be performed simultaneously with the triggering movement and cannot cost more AP than it.

\textbf{Flanking:} Whenever two characters cannot fit completely within a radius of 180 degrees relative to a target, the target must focus on one target at a time. Use 3 surrounding hexes for a hex grid. Switching focus costs 1 AP and is necessary in order to react to the target. If the target of the attack moves as a part of its defense, like with evasive jump, the 1 AP cost is waived as long as the final destination reduces the angle back to 180 degrees or lower.

\textbf{High Ground:} This happens when one character is standing on terrain around 1m higher than its surroundings. The reach is reduced by 1m unless the person on the lower ground targets the legs. The person on the high ground suffers the same penalty to hit the legs of someone on the lower ground. Both receive a +2 bonus to their melee defense against each other. Only long weapons can be used when the height difference is of 2m and so on. The high ground is a defensive position.

\textbf{Cover:} A character can protect themselves against projectiles and explosion attacks by moving behind something, such as an ally or a wall. Staying behind something acts as a shield, reducing the damage taken by a certain amount. A character has cover if the shortest path from the origin of the explosion or projectile to a destination passes through a space containing a blocking object. Cover blocks vision unless it is transparent. Cover can be partial. In that case, it does not block vision and provides a bonus, usually +3, for reflexes and absorbs damage as if it were a shield in the case of grazes and misses. 

\textbf{Underwater Combat:} Any attack that does not have the precise tag or the grapple tag has its damage reduced by half. Fire, radiant and acid damages are also reduced by 50\%. The range of projectiles is reduced to 3m for small objects (the size of arrows), 5m for medium ones (the size of spears) and 8m for large ones (the size of ballista shots, since larger projectiles decelerate less). All explosions have half the radius, but blunt damage is doubled.

\textbf{Mounted Combat:} Attacking in melee receives a mount-dependent penalty while mounted, usually around -2. Accuracy receives the penalty while the mount is moving, not when standing still. Moving while mounted is considered running to trigger reactions, unless stated otherwise. If the character is forcibly displaced while mounted, they will be dismounted.

The only defenses against melee attacks allowed during mounted combat are block and redirect.

\skill{Strength}{STR-10}{ Used to perform feats of strength, such as pushing, dragging, pulling, and lifting things. An active Strength check costs 2 AP + 1 STA. Strength can also be used passively to resist being crushed or moved.

\textbf{Push:} Can be used during a grapple. Make a strength check opposed by the opponent's strength or balance. The defender must also spend 2 AP + 1 STA to resist; otherwise, they receive -5 on the check. A success pushes the enemy 1 space, while a critical knocks them down. Gains +2 when performed during a run.

Multiple characters can participate in a strength conflict. Any skill check made by multiple characters uses the best attributes of any of them, meaning that only one character's strength can be used, plus 4/3/2/1 for each additional character. All characters involved must spend the necessary AP. Smaller characters add a maximum of 2. 

\textbf{Interpreting strength:} The basic rule of strength is that you become 3.33x stronger for every 5 points in this skill, following the VM from the size categories. A STR 10 human can deadlift around 100kg. }

\textbf{Impact from objects:} Objects fall and fly all the time. Here's a table relating object weight and speed and giving the amount of damage it is supposed to deal if it hits a character or another object. This damage is always blunt and its used for non weapon objects, since the speed is relatively slow. 

This table assumes a solid object hitting the target full on. These numbers should be lower for softer materials, like a body, rotten wood or any other deformable object. The suggestion is to lower them by 75\% in the mentioned cases.

\begin{tabular}{| c c c c|}
    \hline &&& \\
    Speed & <5m/s & >5m/s & >12m/s \\
    \hline &&& \\
    0.5kg & 2 & 3 & 6\\
    
    1.5kg & 4 & 6  & 12 \\
    5kg & 6 & 9 & 18 \\
    15kg & 8 & 12 & 24 \\
    
    50kg & 12 & 18 & 36\\
    150kg & 16 & 24 & 48\\
    
    500kg & 24 & 36  & 72 \\
    1500kg & 32 & 48  & 96 \\
    5000kg & 48 & 64  & 124 \\
    
    \hline
\end{tabular}


\textbf{Damage ceiling for impacts:} Strength is used as a ceiling to how much damage can be taken from low velocity projectiles. The maximum amount of blunt damage that can be taken from impact with a projectile is STR x DM x speed tier. Ex: for slow objects under 5 m/s, the ceiling is equal to 5xDM. For medium speed it is 10x STR and for fast ones it is 20x DM. The damage ceiling is applied before reducing damage for soft materials. 

This damage can be defended against and is subject to modifications from attacks, as if they were melee or ranged attacks.

An elephant bumping into a character at low speed will deal, at most, 12 damage.

\textbf{Crushing and being crushed:} Part of the damage a character can take is not from the impact itself, but from the structural force it imposes on the target. This check is called when a target is pinned against a surface, like when an object falls on top of another or in a hammer and anvil scenario. In that case, a Strength check is called to resist being injured or incapacitated by a heavy object. The check is statis, not rolled and costs no AP or STA and has a TN that increases or decreases by 5 by each weight class in the table above. TN 0 is for a 50kg object at low speeds. The TN increases or decreases by 5 for every speed category. A crit does nothing. A hit interrupts, but deals no damage. A graze knocks prone, interrupts and deals a light wound. A miss deals a serious injury, knocks prone and stuns for 5 AP. The injuries can be modified by the rules of locational damage.

\subsection{Stealth and Subterfuge}


\skill{Stealth}{ - }{ Skill to remain unnoticed. The stealth skill is used against stealth events in the environment, which trigger cunning tests for those who will potentially detect.

\textbf{Stealth test:} When a stealth event occurs, the character makes a stealth check against its DL. The result will trigger a cunning test to any characters capable of noticing the event. The DL of the cunning test has a base DL, based on the triggering event. That value is modified by the result of the stealth test. A crit adds +3 to the DL, a hit does not modify, a graze gives a -3 penalty and a miss gives a -5. Crits in stealth tests against passive senses give a +5 bonus to the DL and another +1 for the value that the stealth check goes over the crit threshold. 

Optionally, the GM could roll the dice for the player and not tell them the result. This way, they have to interpret whether they are doing well of not with their stealth, instead of just relaxing and assuming that they won't be seen because they rolled well.

\textbf{Stealth events:} are necessary for a character to remain undetected to another. Without one, characters are automatically detected by each other's active senses. Think about it like an environmental condition that makes stealth possible. A stealth event's DL targets a specific sense, which is the best available sense to use to detect someone in the situation. Normally, the senses to be used are vision and hearing. 

The two big categories of stealth events are: hiding and sneaking. Hiding depends on line of sight and camouflage. Sneaking depends on making no noise when moving, which depends on what the floor is and what is stepping on it.

Some common stealth events and their estimated stealth and cunning DLs are in the following list.}

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item camouflage: stealth DL 0, cunning DL 2 vision
    \item hiding in woods: stealth DL 5, cunning DL 2 vision
    \item hiding in plain sight: stealth DL 8, cunning DL -5 vision
    \item sneaking in the grass: stealth DL 5, cunning DL 0 hearing
    \item sneaking in squeaking floorboards: stealth DL 8, cunning DL -2 hearing
\end{itemize}

\textbf{Remaking Stealth checks:} The stealth check is done once per round in combat, at the beginning of the round, or once per scene outside combat. A scene only changes if the conditions for stealth change significantly. That is a subjective call, but it is recommended that new rolls are not repeated for every action in stealth.

\textbf{Event DL modifier:} Some actions will make the stealth test more difficult. If one of these aggravating events happen within the same scene or round, the last test score is reutilized, adding the new penalty.

Here are the stealth DL modifications:
\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item attacking in melee: +3 hearing and vision
    \item careful movement and crawling: 0 hearing and vision
    \item running: +5 hearing and vision
    \item ranged attacks: +2 to vision
    \item basic movement: +2 hearing and vision
\end{itemize}

\skill{Cunning}{Detection}{ Cunning is used to determine priority when competing to take a combat turn first. It is also used to detect things in stealth or just for looking for things in general.

\textbf{Detection:} Detection uses one of the senses. It uses the cunning skill passively, without adding dice rolls, against the DL of stealth events. On a failure, nothing is noticed. On a graze, reactions to the stealthy characters are allowed at a -3 penalty, but no actions are allowed to be taken against them. A hit allows acting and reacting normally and identifies the target. A critical success allows taking initiative and have your turn immediately.

\textbf{Senses:} The usable senses are vision, hearing and smell. Senses can be active or passive. Active senses are used to keep the detector in alarmed state, meaning that they don't forget your position, even if you move. Passive senses can detect position, but they can't track it or keep the character in alarmed state.

\textbf{Vision:} is an active sense. Stealth event DLs gain +2 from distance per distance band. Vision gets blocked by cover and requires light. The field of vision covers 180 degrees, which is defined in the grid as within the arc defined by the directions of 3 adjacent hexes around a character. In the alarmed state, it covers 360 degrees.

\textbf{Hearing:} is a passive sense. Stealth gains +3 from each distance band and cannot be used to search actively. Hearing is omnidirectoinal.

\textbf{Smell:} is a passive sense. Stealth gains +3 from each distance band. It requires a smell trail to trigger the sense. Humans have -5 to this sense. Sniffing something from very close gives +5 to this test.

\textbf{Touch:} is an active sense. Anything that touches a character is immediately detected and the character becomes alarmed. Projectiles that hit the character have their direction identified, not necessarily their source. Fighting with touch only is only possible while grappling, and takes -2 to Strike and Defend.

\textbf{Distance bands:} The DL from stealth events to cunning is penalized by distance. The magnitude of the modification depends on the specific sense used to detect.
\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item same room (within 10m) - no penalty
    \item at range (within 50m)
    \item far range(within 150 m)
    \item far (within 500 m)
    \item horizon 
\end{itemize}

\textbf{Alarm:} When a character detects the other, they are alarmed to their presence and will not forget their position until they become undetectable to all active senses. Being alarmed is necessary to be able to target something or react to it.

\textbf{Detected and Undetected status} To be considered undetected, a character must be out of the field of detection of all active senses of another character at the beginning of a round. The alarmed state does not wear off until the field of active senses is exited. Each character has their own list of detected characters. 

\textbf{Blind Fighting:} if there are no active senses capable of detecting someone (not event touch), but a hit or crit were achieved in the detection check, it is possible to react or target them at -5 penalty.

\textbf{Darkness and Night:} In a situation of darkness, the field of vision is determined by the light source. If light source is at a distance, consider that the field around it is visible.

\textbf{Looking for objects:} When trying to decide whether characters find an object, the object is found with a hit against the object's TN. If cunning is equal or higher than TN, they find it. 

\textbf{Focusing attention:} This can be used to get +3 to the detection test when looking attentively at one specific thing. It can be used when a player says they want to look at a specific place to see better detail or when someone points out a direction for others to focus. 

\textbf{Investigation:} When a character is looking for something that requires specific knowledge background, they use the value of cunning as skill, but they can only find the object if they fulfill have a knowledge requirement.

\textbf{Analyze Opponent:} Cunning can also be used during a fight to assess an enemy. Spend 3 AP to make a cunning check against an enemy within 30m to try to obtain relevant information about them from visible equipment or physical characteristics. On a critical success, learn important information for combat, such as the creature's vulnerable areas, dangerous items they carry, or special abilities they may have. On a success, obtain a vague reference to a dangerous item or the opponent's power level. Grazes only provide obvious information, and failures provide nothing.
}

\skill{Prestidigitation}{ DEX }{ Represents a character's manual dexterity. It can be used to attempt to steal an object without being noticed, pick a lock, write neatly, conceal small objects, perform delicate operations with the hands, and other manual skills.

\textbf{Standard Action:} Consider that 1 AP is equal to half a second and estimate how long an action should take. In general, taking an item from a belt or strap around the armor should take about 3 AP, while opening a backpack and taking something costs about 5 seconds. The AP cost of a standard action is 3 AP.

% \textbf{Conceal Objects:} Used as a standard action of defense against detection tests to identify objects carried by the character. Successes detect and criticals identify the object. On grazes and failures, nothing is noticed.

\textbf{Theft:} A standard action can be used to steal an item from someone's body or place an item on them. This is done by having one hand free and using a standard action to perform a Prestidigitation skill test against the opponent's cunning +5x the size category of the item (relative to who must detect it). On criticals, it is stolen without being noticed. On successes, the character can choose between successfully stealing or failing without drawing attention. On grazes and failures, attention is drawn and it fails. Stealing an item from stealth gives a bonus of +5.

\textbf{Pick Lock:} These tests require knowledge of mechanics and receive -5 for each level below the requirement. On criticals, it takes 5 AP. On successes, between 10 seconds and 1 minute. On a graze, it takes between 10 minutes and 1 hour, and on a failure, it is impossible to open.

When Prestidigitation reaches 5, the cost of the standard action becomes 2. If it reaches 10, it becomes 1 AP. If Prestidigitation is less than -5, they cost 4 AP.
}

\section{Spells} 

\subsection{Casting Spells}

\skill{Sorcery}{Knowledge+Sorcery}{ Spells are special abilities that require a sorcery test to be used. The value of the sorcery skill is the sum of the sorcery proficiency and the knowledge related to the spell, but the test works the same for every spell}

Casting a spell costs AP and STA during combat or exhaustion during exploration. To cast any spell, make a sorcery test against the \textbf{Spell Casting DC} to determine if the spell was successfully cast or not. In case of failure, it fails and the AP and STA are lost. 

\textbf{Concentration:} All sorcery tests require concentration. The only other action that can be performed while concentrating is careful movement. Being interrupted causes the loss of concentration. Sometimes, a character's AP may not be enough to cover the action's requirements. In that case, they can maintain concentration and continue acting in the next turn. Losing concentration causes all AP and STA spent to be lost. The character must interrupt concentration to be able to actively defend themselves.

\textbf{Sustained Spells:} are actions that can be performed continuously while concentration is maintained. Their effect ends when the caster stops concentrating on it. For these spells, the sorcery skill test needs to be redone periodically, as indicated in the spell. A failure in the sorcery test interrupts the sustained magic.

\textbf{Charged:} can be charged into something and released when a trigger is activated. If the spell is not used within the duration of the charge, it dissipates. 

\textbf{Delayed Spells:} Any spell can be delayed, meaning that the casting can be completed, but the release is postponed. The character must concentrate on the spell until the release, which can be triggered by a previously prepared action. If a trigger does not occur, release it in the next turn. 

\textbf{Requirements:} Some spells may even be impossible to cast if the minimum conditions are not met. Making it rain in the desert is impossible if there is no water in the sky.

\textbf{Stealth and Sorcery:} Spells are imperceptible before they have their effect, unless a skill is used to reveal them. Once their effect occurs, they can be identified with a test of the skill used to cast the spell.

\subsection{Learning Spells}

Spells are trained in the same way as skills, but using the knowledge related to the spell as talent. Each spell costs 4+spell level XP.

They can also be learned with conviction XP or with devotion. In this case, they use their respective levels as talent.

\subsection{Types of Spells and Modifications}

Enhancements are modifications that casters can make to their spells to make them more powerful or better suited to their needs. They are applied after making the sorcery test, but cannot cause the test to graze or fail.

\textbf{Accelerate Spell:} Increases the sorcery DC by 3 to Accelerate a spell for 1 AP. It can be used multiple times to reduce the sorcery time to up to half the original value. It is possible to accelerate spells to avoid opportunity attacks.\\

\textbf{Extend Spell:} Increases the sorcery DC by 2 to increase the range of a spell according to the MM.

\textbf{Enhance Spell:} Increases the sorcery DC by 4 to improve spells in a special way described in the spell.  \\

\textbf{Amplify Spell:} Increases the sorcery DC by 4 to amplify spells. It can be used when the spell has its effects multiplied by DM, SM, MM, and VM. The number of times it can be increased depends on environmental conditions and the material used. The amount of material or charges consumed is multiplied by the VM. Increase Spell allows it to apply to larger versions of the items. \\

\textbf{Extend Charge Duration:} Increases the sorcery DC by 2 to increase the charge duration of charged spells by +50\%, then +100\%, then +200\%, +400\%, and so on in progression. \\

\textbf{Efficient Spell:} Increases the sorcery DC by 4 to reduce the exhaustion or STA cost of a spell by half. \\

The types of spells are:

\begin{itemize}[noitemsep]
\item Environmental: The spell alters the environment.
\item Effect: Spells that directly affect the target and have various effects. They have customized enhancements.
\item Psychic: Spells that directly affect the target and have psychic effects. They use the Charisma proficiency as the Difficulty Level (DL). Enhancing increases the DC by 2.
\item Projectile: Spells that function as projectiles, using the shooting skill as the DC.
\item Touch: Spells that function as strikes, using the striking skill as the DC. They do not trigger opportunity attacks.
\end{itemize}



%----------------

\end{multicols}