\chapter{Character Model}

\begin{multicols}{2}

\section{Character Skills and Actions}

\textbf{Character Actions}: All character actions in the game involve skills in some way. They represent the character's ability to do things. The skill values are added to the dice roll to determine the chance of success for the actions.

\textbf{Action Points (AP):} Action Points are a resource used to represent time. Actions cost AP to perform in situations where time is an important resource. The amount of AP each character receives at the beginning of each round is 6.

\textbf{Skills and Proficiencies}: Skill values are determined by the sum of the skill bonus and the proficiency bonus. Each skill bonus only influences a single skill and can be trained individually. Proficiency bonuses, on the other hand, can affect multiple skills. The default values for all skills and proficiencies is 0.

There are 27 skills, as described in the table:

\end{multicols}

\begin{center}
\begin{tabular}{| c c c c|}
\hline &&& \\
Skill & Proficiency & Talent & Trainable \\
\hline &&& \\
Health & Constitution & Constitution & no\\

Strike & Melee & Dexterity  & yes \\
Defend & Melee & Dexterity & yes \\
Grapple & STR-10 & Dexterity & yes \\

Accuracy & Ranged & Dexterity & yes\\
Reflex & Ranged + Detection & Dexterity & no\\

Balance & AGI-10 & Dexterity  & yes \\
Climb & AGI-10 & Dexterity  & yes \\
Swim & AGI-10 & Dexterity  & yes \\
Strength & STR-10 & Dexterity & yes \\

Prestidigitation & Dexterity & Dexterity & yes \\
Stealth & 0 & Dexterity  & yes \\

Cunning & Detection & Intellect  & yes \\
Exploration & Detection & Intellect  & knowledge \\
Knowledge & 2x knowledge or 2x INT & Intellect  & knowledge \\

Combustion & Sorcery & Intellect & yes \\
Electromagnetism & Sorcery & Intellect & yes \\
Entropy & Sorcery & Intellect & yes \\
Radiation & Sorcery & Intellect & yes \\
Animancy & Sorcery & Intellect & yes \\
Telepathy & Sorcery & Intellect & yes \\
Biomancy & Sorcery & Intellect & yes \\
Devotion & Devotion & Spirit & yes \\

Insight & Charisma & Intellect & yes \\
Deception & Charisma & Intellect & yes \\
Persuasion & Charisma & Intellect & yes \\
Will & SPI & Spirit & no \\

\hline
\end{tabular}
\end{center}

\begin{multicols}{2}
    

Skills marked as trainable can be improved individually. The ones that are not can only get bonuses from their proficiency.

\subsection{Proficiencies:}

The trainable proficiencies are as follows:

\textbf{Melee (DEX)}: General skill in melee combat. Uses DEX as talent. \\
\textbf{Ranged Combat (DEX)}: General skill in ranged combat. Uses DEX as talent. \\
\textbf{Detection (INT)}: The character's ability to pay attention to what matters and not get distracted. Uses INT as talent.\\
\textbf{Charisma (INT)}: General skill of communication, both verbal and non verbal. Uses INT as talent. \\
\textbf{Sorcery (SPI)}: General skill in casting spells. Uses SPI as talent. \\
\textbf{Devotion (SPI)}: Level of alignment with a deity. Uses SPI as talent.\\
\textbf{Conviction (SPI)}: Alignment of worldview and instinct. Uses SPI as talent. \\

\subsection{Levels}

The expected levels for skills and proficiencies range from 0 to 5, but can be increased even further.

\textbf{Level 0, Novice:} Knows only what is common sense\\
\textbf{Level 1, Apprentice:} Can do the basics\\
\textbf{Level 2, Amateur:} Masters the basics and has some advanced knowledge\\
\textbf{Level 3, Professional:} Works at a professional level\\
\textbf{Level 4, Expert:} Capable of reproducing almost anything in the field \\
\textbf{Level 5, Master:} Capable of producing results of the highest level.\\
\textbf{Level 6, Grandmaster:} Is pushing the boundaries of knowledge in their era.\\

\section{Talent and Learning}

\textbf{Gaining Experience (XP):} Improving skills and proficiencies and learning abilities requires training during free time, which yields experience based on training time. A regular training session for skills and proficiencies takes 10 days for 1 XP and costs a certain amount of money based on the difficulty level of what one wants to learn.

\textbf{Talent and XP Costs}: The cost of XP to improve a proficiency is equal to $4+4*n$, where n is the proficiency level one wishes to achieve. If the talent is lower than the level one wants to reach, this value is multiplied by $2^x$, where x is the difference between n and the talent, i.e., n-talent. The cost to increase a skill is equal to half the cost to increase the proficiency.

A character has 4 talents that represent their potential for growth.

\textbf{Constitution (CON)}: Represents the character's general health and ability to adapt physically.

\textbf{Dexterity (DEX)}: Represents the character's motor skills and reaction time. Includes all skills related to athletics and combat. These are well explained in the chapter on combat.

\textbf{Intellect (INT)}: Represents memory, logical reasoning, and attention management. Includes all knowledge, most social skills, exploration skills, and cunning in combat. 

\textbf{Spirit (SPI)}: Represents the character's spiritual powers. It is the primary attribute used for spellcasting, as well as abilities derived from conviction and devotions.

\textbf{Improving Talents:} During character creation, the player must choose the character's talents in order of priority. After that, talents can only be increased by spending karma. The karma cost to increase a talent is equal to $4+4*n$, where n is the talent level one wishes to acquire. The maximum talent value is 5.

\textbf{Level Limits:} The values of talent and proficiency range from 0 to 5, and cannot be increased beyond that. Skills can be increased indefinitely, but they become exponentially more expensive since the maximum talent is 5.

\subsection{Karma}

\textbf{Karma:} is used to increase talents. It is received upon completing missions and their secondary objectives, as well as being rewarded for good gameplay.

The Master can reward players for loyalty to their character, creative use of the system, innovative problem-solving, and participation in dramatic events.

\textbf{Loyalty to the character:} This means the player acts in ways that the character would, rather than what the player thinks is best.

\textbf{Creative use of the system:} This is an extra reward for finding new ways to be effective with your character using the system. Exploring weaknesses in enemies, winning a battle without taking damage, and resolving a violent situation diplomatically are examples of actions that can generate karma.

\textbf{Participation in dramatic events:} Speeches, discussions between characters, exploration, and conversations that show curiosity about the world, demonstration of the character's ethical and moral flaws, and comedic moments.

Karma can also be subtracted as a penalty for acting against the principles of any of these rewards.

\section{Constitution}

Constitution governs the enhancement of attributes, health, and also mutation abilities.

\textbf{Health:} The character's ability to heal wounds and resist toxins. The use of health is described in the chapter on survival.

\subsection{Attributes}

Attributes represent the physical characteristics of the character. They are divided into Strength (STR), Agility (AGI), and Stamina (STA). The common values for attributes for healthy young humans are 10.

\textbf{Improving attributes:} The XP cost to improve attributes is equal to the target attribute value divided by 2. However, if the sum of the 3 attributes plus the target value exceeds $40+4*CON-4*SM$, the cost is doubled. For every 3 points higher, it triples, and so on for every 3 points.

\textbf{Changing attributes:} It is possible to lose 1 point in one attribute to increase another, but this costs 2 XP.

\subsubsection{STR}

Represents the weight and muscular strength of a character. Characters with higher STR can hit harder, resist damage better, and perform more impressive feats of strength.

\textbf{Natural Toughness (TGH natural), Natural Resistance (RES natural), and Natural Insulation (INS natural)} define how much damage a character can withstand before receiving cuts, impact, and burn injuries, respectively. Their default value for all three is 0.5xSTR xDM.

\subsubsection{AGI}

Represents the character's athletic ability. Agility contributes to proficiency in swimming, climbing, and balance. Additionally, it defines movement speeds and gives bonuses to some combat moves.

\subsubsection{STA}

Represents the total breath and physical energy of the character. It is used to define tolerance to exhaustion. STA is used as a resource for all types of actions that require physical effort and also for bursts of action.

\textbf{Rest:} is an action that costs 4 AP and recovers STA by an amount equal to STA/4.

\subsection{Size}

Size is a characteristic that defines how much space a creature occupies on the grid and modifies the physical qualities of characters, weapons, and armor.

Consider that the weight of a creature increases 10 times every 2 categories or 3 times for each category. A human is a creature of size 3 and weighs about 50kg. Therefore, a creature of size 1 would be a cat weighing 5kg, and a horse would be a creature of size 5, as it weighs about 500kg.

A creature of size 3 in human form occupies 1 square or hexagon. For every 2 size categories, a character on a square grid will increase their dimensions by 1 square, and a hexagonal grid will occupy 3 spaces (placing the token in the middle of 3), then 7, then 19 (with the token in the central hexagon and expanding outward by 1 hexagon).

Two creatures of size 3 can occupy the same hexagon, but both receive the grappled condition while doing so. The number of creatures increases by 1 for each size below 3, and the penalty only applies when the limit is reached. Larger creatures cannot occupy the same central space, but they can share the edge hexagons, also receiving the grappled condition. The condition does not apply to the larger creature if its size is 2 greater than that of the smaller.

\textbf{Increasing a creature:} The strength of a creature increases with the cross-sectional area of a muscle, while weight increases with volume. Therefore, when increasing the size of a creature without changing its shape, it receives a penalty of -1 STR and -1 natural AGI.

\textbf{Standard Deflection (SD):} is the standard difficulty to hit a target, which is based solely on its size and movement. The larger the target, the easier it is to hit. The standard deflection is equal to -5-1xSM for stationary targets and -2-1xSM for moving targets. It is impossible for any type of penalty to reduce a character's deflection to a value lower than the SD.

\textbf{Damage Multiplier (DM):} multiplies the HP of a character, the character's damage tolerances, and their armor, weapon damage, abilities, and spells.

\textbf{Skill Modifier (SM):} is a modifier added to or subtracted from the character's values, in the proportions shown in the table.

\begin{center}
\begin{tabular}{||c c ||}
\hline
Skill & Modifier   \\
\hline
Grapple & +5x \\ 
Strength & +5x \\ 
Reflex & -1x \\ 
Armor Cover & -1x \\ 
Weapon Precision & -1x \\ 
SD & -1x \\ 
Stealth & -3x \\
Climbing & -2x \\ 
\hline
\end{tabular}
\end{center}

\textbf{Movement Multiplier (MM):} multiplies all movement speeds and the range of all weapons.

\bigskip

\begin{center}
\begin{tabular}{||c c c c c c c c||}
\hline
Size & 1 & 2 & 3 & 4 & 5 & 6 & 7 \\
\hline
DM & 0.5 & 0.75 & 1 & 1.5 & 2 & 3 & 4 \\
SM & -2 & -1 & 0 & +1 & +2 & +3 & +4 \\
MM & 0.5 & 1 & 1 & 1.5 & 1.5 & 2 & 2.5 \\ 
VM & 0.1 & 0.33 & 1 & 3.33 & 10 & 33.3 & 100 \\ 
\hline
\end{tabular}
\end{center}


\section{Races and Age}

\subsection{Races}
Each playable race comes with its own bonuses to attributes and talents. The table below shows the playable races:


\begin{center}
\small
\noindent
\hspace*{-1cm}
\begin{tabular}{||c c c c c c c c||}
\hline
Race & STR & AGI & STA & INT & SPI & DEX & CON \\ [1em]
\hline\hline
Surbit & 1 & 2 & 2 & 0 & 0 & 1 & 1 \\[0.5em]
\hline
Dwarf & 1 & -1 & 0 & 1 & 1 & 0 & 0 \\[0.5em]
\hline
Human & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\[0.5em]
\hline
Halkir & 0 & 2 & 1 & 0 & 0 & 1 & 0 \\ [0.5em]
\hline
Erthal & 2 & 0 & 1 & 0 & 0 & 0 & 1 \\[0.5em]
\hline
\end{tabular}
\end{center}

\textbf{\large Surbita}
\begin{itemize}
\setlength\itemsep{0em}
\setlength{\parskip}{0em}
\item Small-sized creature
\item Has the Fur ability
\item Has the Bear Nose ability
\item Has the Tail I ability
\item Has the Claws ability
\end{itemize}

\textbf{\large Dwarves}
\begin{itemize}
\setlength\itemsep{0em}
\setlength{\parskip}{0em}
\item Has the Bat Ears ability
\item Has the Tireless ability
\end{itemize}

\textbf{\large Humans}
\begin{itemize}
\setlength\itemsep{0em}
\setlength{\parskip}{0em}
\item Can choose 2 talents to have +1.
\item Can choose one from Tireless, Large Hands, and Beautiful.
\end{itemize}

\textbf{\large Erthal}
\begin{itemize}
\setlength\itemsep{0em}
\setlength{\parskip}{0em}
\item Has the Large Hands ability
\item Has the Regenerative Reservoir ability
\end{itemize}

\textbf{\large Halkir}
\begin{itemize}
\setlength\itemsep{0em}
\setlength{\parskip}{0em}
\item Has the Beautiful ability
\item Has the Infrared Vision ability
\end{itemize}


\subsection{Aging}

A character's age influences their attributes and ability to learn new things. A character can be created as a child, a teenager, an adult, or older. Each of these ages has its characteristics.

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item Child: -2 STR, talents start at 0 but cost half the karma. Considered a size 2 creature.
    \item Teenager: -2 STR, +1 to talents.
    \item Adult: Standard.
    \item Aging: Receives -1 to CON and also to all attributes at each stage.
\end{itemize}

The time required to advance in the age category can vary between races.

\begin{center}
\normalsize
\begin{tabular}{| c c |}
\hline
Stage & Human \\ [1em]
\hline
Child & 8-13 \\ [0.5em]
Teenager & 14-18 \\[0.5em]
Adult & 19-39 \\[0.5em]
Mature & 40-59 \\[0.5em]
Old & 60-79 \\[0.5em]
Very old & 80-death \\[0.5em]
\hline
\end{tabular}
\end{center}


\section{Intellect}

Intellect measures the logical capacity, planning, and decision-making abilities of characters. The skills governed by intellect involve exploration, knowledge, sorcery, and cunning. 

\textbf{Exploration:} It is used in survival situations to determine the ability to travel without getting lost and to find things in varied terrains. The exploration skill is based on the sum of detection proficiency and specific knowledge of the area.

\textbf{Sorcery:} Uses sorcery proficiency added to a related knowledge to cast spells. More details are in the combat section.

\textbf{Cunning:} It is the ability to make good decisions quickly. It is the sum of detection and cunning skill. More details are in the combat section.

\subsection{Knowledge}

Knowledge is like skills, but it is generally used outside of combat and has various utilities related to magic, exploration, and measuring the character's intelligence.

\textbf{Knowledge as a skill:} The purest way to use knowledge is as a skill:

\skill{Knowledge}{ 2x Knowledge or 2x INT}{
It is equivalent to asking your character if they should know or understand something, even when the player does not know that thing.

You use 2x the relevant knowledge as the skill value. If the necessary facts for deduction are already known, you can use 2x intellect as the value.

Whenever a player claims that the character knows something, the Master can ask for a knowledge check to determine if the character really knows it.

To use this skill, make a knowledge check against an arbitrary difficulty, representing the challenge of reaching a good conclusion from the available information. Critical successes mean that the character can make the best deduction from the information they have, even determining what information is missing to make a better deduction and remembering everything they know. Successes mean they have a good guess and remember most of what they know. Misses mean that you don’t know or can’t remember, but you might know where to look. Failures mean that you reach an incorrect conclusion or a different one from the correct one, or you can’t remember anything.
}

% Skill used to learn and remember facts. It can be used to actively seek and acquire information of any kind from books, maps, artifacts, and similar sources. When using it, make a skill check to learn as much as possible from the available material, considering how good that material is and how much time you have to study.

\textbf{Knowledge as a talent:} Knowledge can be used as a talent to learn spells. Spells have a knowledge requirement indicating the difficulty.

The formal areas of knowledge in which the character can be trained are:

\begin{itemize}
    \item Alchemy: Knowledge about potions, poisons, oils, and other chemicals. 
    \item Armory: Knowledge about the manufacture and maintenance of weapons and armor. 
    \item Animancy: Knowledge about the soul and its effects on the world. What it does in the body, how it manifests and affects the environment, how to manipulate it, etc. Necessary for magic.
    \item Architecture: Knowledge about structures, their common forms, how to build or destroy them. Tunnels, bridges, buildings.
    % \item Botany: Knowledge about plants, their location, shape, magical properties, poisons, fruit-bearing capacity, ecology, etc.
    
    \item Combustion: Knowledge about combustible materials and their properties. Necessary for magic.
    \item Electromagnetism: Knowledge about electricity and magnetism. Necessary for magic.
    \item Entropy: Knowledge about the flow of energy and its manipulation. Necessary for magic.
    \item Geography: Knowledge about maps, geology, and geography. Allows for map-making.
    \item Geopolitics: Knowledge about factions, noble families, military history, political ideologies. 
    \item Mechanics: Knowledge about mechanisms. Includes locks, siege weapons, 
    \item Radiation: Knowledge about light and radioactivity. Necessary for magic.
    \item Biomancy: Knowledge about controlling living structures. Necessary for magic.
    
    % \item Medicine: Knowledge about diseases, anatomy, and biological processes. Necessary for magic.
    \item Survival: This is the knowledge used in the action of exploration and preparation. It includes actions like cooking, butchering animals, preparing campfires, making natural shelters, and crafting simple objects from materials found in nature.
    \item Telepathy: Knowledge about the transmission of information through magical means. Necessary for magic.
    % \item Zoology: Knowledge about animals, their anatomy, weaknesses, ecological roles, habits, etc.
\end{itemize}
\noindent

\textbf{Learning Knowledge}: The knowledge listed in the areas of knowledge is learned exactly the same way as any skill, using INT as the talent.

\subsection{Informal Knowledge}

Informal knowledge is the result of the character's experiences and can be acquired by engaging in any activity. The categories can be diverse, created to represent the character's experiences.

\textbf{Varying XP cost:} The amount of XP required to level up can also be arbitrary, depending on the difficulty of learning the knowledge. For example, knowing all the relevant people in a town of 400 inhabitants is much easier than learning a new language. INT is used as a talent for skills normally.

Examples of varied knowledge:

\textbf{Location - place:} Knowledge about the people in the town and their local customs. \\
\textbf{Survival - region:} Knowledge about a specific region, its roads and paths, food sources, water sources, natural disasters, economic resources, geographical landmarks, etc. This type of knowledge is often used alongside survival actions like traveling and searching. \\
\textbf{Language:} Knowledge about a specific language. Having level 4 in a language is necessary for fluent speech. \\
\textbf{Art - type:} Knowledge about a specific form of art, for example, music, its greatest artists, and works. \\
\textbf{Faction - name:} Knowledge about the people of a faction, their history, politics, wealth, alliances, etiquette, etc. \\
\textbf{Religion - name:} Knowledge about the gods of a religion and their domains. \\

\textbf{Receiving XP:} Downtime activities generate experience in knowledge. The character receives XP in knowledge simply by having access to related information. The amount generated will depend on the character's dedication and ability to learn. For example, if the character is in a town but spends all their time learning sorcery in a tower, they do not gain local knowledge. If they stay in the tavern meeting new people and celebrating, but do not know the language, they gain a lot of knowledge about the language, but less about the local area.

\textbf{Reward for specific feats:} XP can also be generated by specific events. For example, experiencing multiple exploration events in a certain region can generate survival XP in that region.

\subsection{Social}

Characters have skills that define how proficient they are in social situations. Charisma is the main proficiency that governs social skills.

There are three skills relevant to social situations. Empathy, Rhetoric and Will.


\skill{ Persuasion }{ Charisma }{ This is used to read character's intentions, understand their desires and emotions, as well their connection to behavior. Charisma can be used to detect when someone is lying, even when the logic makes sense.

Conversely, it can be used to control facial expressions, body language, your accent, to know the appropriate things to say. This is the actor's skill. It is important for someone to make performances, impersonate someone, infiltrate environments without being noticed socially. This is the skill used to tell lies and not be noticed.
    
Empathy is used in spells that involve convincing others and reading their minds. It is a subtle kind of manipulation, which makes people think that an idea was their own.}

\skill{ Insight }{Charisma}{ This is the skill used to understand others' motivations and detect deception. It is used in social interactions to understand others' intentions and desires, as well as to detect when someone is lying, even when the logic makes sense. 

This can also be used to detect a person or creature's emotional state.

NPCs have tags on them that indicate what they value and what they want out of an interaction. Insight can be used to identify these tags to make it easier to manipulate them.}

\skill{ Deception }{Charisma}{ This is the skill used to lie and not be noticed. It is important for someone to make performances, impersonate someone, infiltrate environments without being noticed socially. This is the skill used to tell lies and not be noticed. }

\skill{ Will }{ SPI }{ Will is like the character's presence and its mental fortitude. A character with strong will is more resistant to emotional stress and to magical mind altering effects. In social situations, it is used to intimidate people and to resist being intimidated.
    
It is used in spells that involve causing fear, anger, and conflict in general. People with high Will make strong impressions and can be very intimidating. }

\skill{Rhetoric}{Charisma+INT}{This is the ability to make an argument make sense and to notice inconsistencies in other's arguments. 

It is used in social interactions when trying to convince someone of something logically and also to find inconsistencies in someone else's logic. Rhetoric is resisted by rhetoric. }


% Fatalism: +5 Will \\
% Stoicism: +3 Stress, +3 Will \\
% Hedonism: +5 Charm \\
% Guardian: +3 Stress, +3 Will \\
% Divine Providence: +3 Charm, +3 Will \\
% Ferocity: +5 Stress\\
% Adaptation: +2 Charm, +2 Stress, +2 Will\\
% Domination: +3 Charm, +3 Stress, +3 Will\\

\subsection{Contacts}

\textbf{Contact}: is a trusted character. Contacts can answer sensitive questions, offer better prices, perform favors, and can buy and sell illegal items. The type of contact will define what they do. The types of contacts are described in the chapter on social interactions.


\section{Spirit}

\subsection{Limit Actions:}

Each character must choose a limit action.

\textbf{Coward:} is compelled to use their movement surge to run away and try to hide.

\textbf{Violent:} is compelled to use their fight or utility surge to inflict damage on someone.

\textbf{Tanatosis:} Falls to the ground paralyzed for 1 turn and rests 2 times.

\textbf{Abusive:} is compelled to use their utility surge to insult allies, increasing the test by +1 for the team.

\textbf{Traitor:} is compelled to use their utility surge to negotiate with the opponent. If the target is unable to communicate, they run for 1 turn.


\subsection{Convictions}

Convictions provide access to various buffs, exclusive abilities, and are the only way to increase the character's will, but they impose some restrictions on the character's behavior.

\textbf{Acquiring a conviction:} A character must adopt one worldview conviction and one instinct conviction during character creation. Each has a defined way of gaining XP through the character's actions.

\textbf{Evolving Convictions}: Each conviction has its XP and level tracked individually. SPI is the talent used to determine the experience cost of improving the conviction. Conviction abilities are automatically acquired at each level.

It is possible to lose conviction XP. If the session ends with negative XP, the proficiency loses a level. Each conviction has its condition for XP loss. If the level reaches 0, the conviction is considered renounced.

There are 8 convictions arranged along two axes: Extrospective - Introspective and Order - Chaos. 4 of them are worldview convictions and 4 are instinct convictions. The convictions are detailed in the chapter on skills.

\subsection{Deities and Devotion}

One can choose to be a cleric of a deity. To do this, one must have at least one conviction in common with the deity and acquire the cleric ability.

Being a cleric of a deity grants access to cleric abilities and also allows one to learn all the skills and spells favored by the deity using devotion XP and the devotion level as talent. The Devotion skill can be used to cast spells related to the entity.

Each deity has attitudes that are promoted and disapproved.

\subsubsection{Ateu}

Values hard work and commitment. \\

CONTRACT: Gains the Contract ability. When the contract is made, both characters define a completion condition and a penalty. Gains 1 XP for making a contract, a variable amount of XP for completing the contract, and loses the same amount of XP for failing to comply. The loss can be mitigated by paying a penalty. \\

CANCELLATION: Atheist clerics always know if the parties are alive and if they have upheld the contract, as long as they are within areas of divine reach. The contracted individuals are marked and can be tracked by any Atheist cleric until they pay the penalty or complete the contract. \\

Convictions: Ferocity and Adaptation \\

Abilities: Gas mask, magnetic shield, electric skin \\

Spells: Lamp, Revealing Light, Dazzling Flash, Sharpen, Rust, Repair metal, Start fire, Flame thrower, sustained lightning, activate explosive, alchemical grenade, ghostly sound, Taser, Firestorm, Aurora, freeze water, rain. \\

\subsubsection{Iskar}

Values courage and justice \\

TRIAL: Learns martial spirit and uses devotion instead of animancy as a skill. \\

REDEMPTION: Swears to redeem someone's soul. If able to subjugate them, allows the choice between the effects of mind disintegration as redemption or death. Gains variable XP each time this occurs or half that if the target does not accept and is killed. Loses half the value if entering combat and surrendering. \\

Convictions: Fatalism and Stoicism \\

Abilities: Synesthesia, Spiritual Strike, central intelligence, blind combat, psychoactive assistance, chaotic thorn, \\

\subsubsection{Minala}

Values 

FAITH TEST: Gains healing food. Heals if true and well-intentioned, but poisons and loses nutritional value if not. \\

RESURRECTION: Faithful clerics of Minala can be revived if placed inside one of her temples, following the restrictions of the resurrection ability. \\

Convictions: Divine Providence and Guardian \\

Abilities: \\

Spells: Accelerate Blooming, enhanced sleep, treehouse, control tree, healing touch, Defibrillate, regeneration, Reattach limb\\

\subsubsection{Rivka}

Values diplomacy and intelligence \\

CONFESSION: Gains the ability to alter memories. Receives 1 XP for using it. Loses 4 XP if revealing the secret or if the person realizes that their memory has been altered. Performs a confession ritual in which they see the memories associated with the story. Can alter the person's memory, but must not let them perceive the alteration. \\

DIVINE GUIDE: The cleric receives 1 exhaustion to create a ghost of Rivka. The ghost may help the group or not, depending on its will. \\

Convictions: Hedonism and Domination \\

Abilities: Synesthesia, read surface thoughts, central intelligence, sculpt face, universal communication, spiritual walk. \\

Spells: telepathic link, scare, provoke, calm, visual interference, suggestion, ghostly sound, Command, possession, create fog, rescue, psychoactive assistance, astral travel, torment, anchor soul, and Command. \\

% end --------------------------------------------------------------------------------------------------

\section{Character Creation}

\textbf{Step 1:} Choose the character focus. Wealth, DEX, CON, INT, or SPI. Assign each focus a priority level, from A to E. \\
It is possible to start at an older age and be allowed to select two focuses at A and none at E.

\textbf{Step 2: Race.} Choose a race, which grants racial abilities and bonuses to attributes and skills. At this stage, you also choose the age. 

\textbf{Step 3: Wealth.} The character starts with \\
A: 100 GP, \\
B: 60 GP, \\
C: 30 GP, \\
D: 10 GP, \\
E: 5 GP.\\

\textbf{Step 4: CON.} Characters start with 8 in all attributes and receive XP to increase them depending on the focus. \\
A. CON=2, +60 XP. \\
B. CON=1, +42 XP. \\
C. CON=1, +30 XP.\\
D. CON=0, +24 XP. \\
E. CON=0, +16 XP.\\

\textbf{Step 5: SPI.} Starts with XP to distribute among the proficiencies, skills, and SPI abilities. \\
A: SPI=2, 60 XP,  \\
B: SPI=1, 42 XP,  \\
C: SPI=1, 30 XP, \\
D: SPI=0, 24 XP, \\
E: SPI=0, 16 XP. \\

\textbf{Step 6: INT.} Starts with XP to distribute among the proficiencies, skills, and INT abilities. \\
A: INT=2, 60 XP \\
B: INT=1, 42 XP, \\
C: INT=1, 30 XP, \\
D: INT=0, 24 XP, \\
E: INT=0, 16 XP. \\
All characters start with one language at level 4.\\

\textbf{Step 7: DEX.} Starts with XP to distribute among the proficiencies, skills, and DEX abilities. \\
A: DEX=2, 60 XP. \\
B: DEX=1, 42 XP, \\
C: DEX=1, 30 XP \\
D: DEX=0, 24 XP \\
E: DEX=0, 16 XP.\\

\end{multicols}
