\chapter{Wealth and Equipment}

\begin{multicols}{2}

\section{Weapons Properties}

A weapon is a tool used in combat to enhance the character's abilities. Each weapon or shield has at least one form of attack. Each form has its own properties: \\

\noindent
\begin{itemize}
\setlength\itemsep{0em}
\setlength{\parskip}{0em}
\item \textbf{Grapple:} This attack is used for grappling actions, its force is equal to STR.
\item \textbf{Sweeping:} Allows for Sweeping attacks.
\item \textbf{Braced:} This attack can be a braced attack.
\item \textbf{Throw:} indicates that the throw attack can be used.
\item \textbf{Handle:} This attack represents the handle of the weapon.
\item \textbf{Short, Long I/II:} Short-range attacks can be used in grappling combat but have a range of 1m x MM. Long-range attacks have an extended range of 1m x MM per level and cannot be used in grappling combat.  
\item \textbf{DEF:} Attacks that have DEF can be used to block. 
\item \textbf{STR x:} Strength requirement.
\item \textbf{Hook:} This attack can be used to knock down.
% \item \textbf{BLT/CUT:} indicates the blunt and cut damage. CUT is always a multiple of BLT.
\item \textbf{Small/One/Two hands:} Small weapons give +2 to Strike against oponents involved in the same grapple as oneself. One-handed attacks use only one hand. Two-handed weapons require both hands. If used with one hand, they cost +1 AP for both attack and defense.
\item \textbf{Heavy I/II/III:} Allows heavy attacks. Having a higher degree of heavy allows using any lower degree. Having heavy I-III or similar means that heavy I is minimum.
\item \textbf{Piercing:} Grazes behave like a miss. Divides the amount of IL received from cutting damage by its SHP. Also reduces damage to gear and objects by the same amount.
\item \textbf{Smash:} allows using the smash additional effect.
\item \textbf{Bladed:} allows using the extra cut additional effect and can be sharpened.
\item \textbf{Penetrating:} allows using the penetrating additional effect.
\item \textbf{Precise:} allows performing armor bypass additional effect.
\item \textbf{RES:} indicates RES of that part of the weapon, which is the threshold for gear breakage.
\item \textbf{Fast, UF:} Slow: Force is 2x energy. Fast: can only blocked with a shield. Ultra fast(UF): like fast, but it is automatically penetrating. 
% \item \textbf{Overpen:} Maximum IL damage is 10.
\item \textbf{Draw:} Attack can be performed at the same time as drawing the weapon.
% \item \textbf{Shot:} This can only blocked with a shield.
\item \textbf{Reload:} The weapon requires reloading to attack again.
\item \textbf{Size x:} Defines which size of creature can wield the weapon. Wielding a weapon that is larger than appropriate gives a increases the cost of AP for all attacks by +1, and uses STR-5 in the contribution of STR to damage. This penalty applies to any contribution of STR for any attack, including braced, hooked, and heavy. Wielding a weapon 2 size categories larger is impossible. 
\end{itemize}

Some human-sized versions of weapons are shown in the list. To create a version of a different size, adjust the DM to the appropriate size category. \\

\section{ Weapons}

\subsection{Melee Weapons}

\textbf{ Unarmed:} Corresponds to the hands of a humanoid character. It has punch and grapple as attacks. The punch deals damage based on the character's strength. A punch is unlikely to damage another character wearing armor, but since it is a small weapon with short range, it can land vulnerable hits with some ease. A heavy attack that overcomes armor coverage may be enough to knock someone out. The grapple does not deal damage, but serves both to hold the opponent and perform grappling maneuvers as well as to block some damage.

\textbf{Brass Knuckles:} are a stronger version of the fists that are part of the armor. The character receives penalties for prestidigitation, accuracy, and climbing, but gains more coverage in the armor and has a much stronger and more resilient punch.

\textbf{Rondel:} is a dagger focused on destroying armored opponents. Its penetration is quite high, making opponents wearing a chain shirt or chainmail extremely vulnerable. Being a small weapon with short range, the rondel is ideal for grappling combat. Even when its penetration is insufficient against armor, the rondel, being a precise weapon, can devastate the opponent if it lands a vulnerable attack.

\textbf{Dagger:} are extremely versatile weapons. They can be used as tools or as weapons and are easily concealed. Daggers are also used as defensive tools, making them useful when held in the non-dominant hand.

\textbf{Whip:} is the only light weapon with long range. Its damage and penetration are insufficient to injure any armored opponent, but its great advantage lies in the grapple attack. Keeping a target under control at a distance combined with the ability to drag it towards you can be a great strategic move against agile opponents, leaving them open to braced and opportunity attacks, as well as reducing their ability to defend. There are modified versions of whips, such as the electric whip, which allows electrical abilities to hit grappled opponents.

\end{multicols}

\begin{table}[H]
\scriptsize 
\centering
\renewcommand{\arraystretch}{1.4}
\hspace*{-1cm}
\begin{tabular}{ | c c c c c c c c c |}
\hline
\multicolumn{9}{|c|}{ \textbf{ \Large Small Arms} } \\
\hline
Weapon & RES & Energy & SHP & hand & range & properties & price & size \\
\hline
\textbf{Unarmed} & TGH & 0 & 0 & small & short & grapple, DEF, draw & - & small \\
& STR & 0.5x STR & 0 & small & short & draw & smash &  \\[5pt]
\hline
\textbf{Claws} & 10 & 0.5x STR & 1x & small & short & grapple, DEF & - & small \\[5pt]
\hline
\textbf{Brass Knuckles} & 25 & 0 & 0 & small & short & grapple, DEF & 15 GP & small \\ 
& 25 & 0.5x STR & 1x & one & short & heavy I, smash & &\\ [5pt]
\hline
\textbf{Dagger} & 15 & 8 & 2x & small & short & draw, DEF, piercing, bladed, precise & 3 GP & small\\
& 15 & 8 & 2x & small & STR & draw, precise, throw && \\ [5pt]
\hline
\textbf{Stiletto} & 15 & 6 & 3x & small & short & draw, penetrating, precise, piercing & 5 GP & small\\
& 15 & 0 & 0 & small & short & draw, DEF && \\ [5pt]
\hline
\textbf{Whip} & 10 & 4 & 1x & small & long II & draw & 2 GP & small\\
& 15 & 0/4 & 3 & small & long II & grapple & & \\[5pt]
\hline
\end{tabular}
\end{table}
\begin{multicols}{2}

\textbf{Short Sword:} It is a larger version of a dagger, suitable for public use without drawing attention and easy to draw. Although it is too large to be ideal for grappling combat, it has stronger attacks and offers better defensive potential. It is a good secondary weapon or for urban and enclosed environments. 

A variation of the short sword is the scimitar, which receives +1 to damage on slashing attacks and -1 on thrusts.

\textbf{Rapier:} It is the fastest one-handed long weapon available. They are excellent for duels against lightly armored opponents, as their flurries of strikes are very difficult to defend against, and their range provides a significant advantage over shorter weapons. Rapiers are commonly seen in urban settings and are easily carried in enclosed spaces. They are often used in conjunction with shields or daggers, as these enhance their defensive potential and open opponents up to counterattacks.

\textbf{Short Spear:} The most common weapon on the battlefield. They can be used one-handed alongside shields or two-handed to increase strike speed. They can also be thrown, making them a very versatile weapon. Short spears are very cheap and common, but they are unusual in an urban environment and should be carried in one hand or strapped to the back. Due to being a wooden shaft weapon, the durability of spears can become an issue when used against highly rigid objects or when used to block cutting weapons.

\textbf{Mace:} It is a weapon focused on taking down armored opponents. Its best feature is its heavy strike, which is relatively quick compared to the heavy strikes of other weapons. The durability of the mace is very high when used offensively, but its shaft makes it fragile defensively. Some maces can be made entirely of metal.

\textbf{Battle Axe:} It is a very good option against wooden objects or opponents with low-rigidity armor. Its blade combined with its weight makes it an excellent weapon for destroying shields. The axe is slower than swords, and its attacks can leave the user open to enemy offensives, so it is wise to make each attack count. Its defensive potential is low, as it is a shaft weapon. It is better to use the axe alongside a shield to compensate. The battle axe has a spike on the back, which is a better option against opponents whose armor the blade cannot penetrate.  

\end{multicols}
\begin{table}[H]
\scriptsize 
\centering
\renewcommand{\arraystretch}{1.4}
\hspace*{-1cm}
\begin{tabular}{ | c c c c c c c c c |}
\hline
\multicolumn{9}{|c|}{ \textbf{ \Large One-Handed Weapons} } \\
\hline
Weapon & RES & Energy & SHP & hand & range & properties & price & size \\
\hline
\textbf{Short Sword}& 20 & 8 & 2x & one & short & precise, draw & 8 GP & medium\\
& 20 & 10 & 1.5x & one & short & sweep, draw, DEF, heavy I, bladed  && \\ [5pt]
\hline
\textbf{Rapier} & 15 & 8 & 2x & one & long I & draw, precise, piercing & 10 GP & medium\\
 & 15 & 0 & 0 & one & long I & DEF & & \\[5pt]
\hline
\textbf{Short Spear} & 20 & 8 & 2x & two & long I & braced, precise, piercing, heavy I & 1 SP & large \\
& 15 & STR & 1x & two & short & DEF, shaft, grapple &  & \\
& 20 &  8  & 2x & one & 2xSTR & throw, precise, heavy I-I && \\[5pt]
\hline
\textbf{Mace} & 50 & 12 & 1x & one & short & heavy I-II, smash & 4 GP & medium\\
\textbf & 15 & 0 & 2 & one & short & DEF, shaft &  & \\[5pt]
\hline
\textbf{Battle Axe}& 25 & 10 & 1.5x & one & short & heavy I-II, bladed & 5 GP & medium\\
& 15 & 0 & 2 & one & short & DEF && \\ [5pt]
\hline
\end{tabular}
\end{table}

\begin{multicols}{2}

\textbf{Bardiche:} is a large axe with a long blade at the tip. It has much in common with the battle axe, but it is used with two hands, which greatly increases its destructive potential. Additionally, the size of the blade allows it to deliver sweeping attacks, hitting multiple opponents at once. The bardiche does not have a good option for attacking opponents with high rigidity armor. These characteristics make it a relatively specialized weapon. 

\textbf{Longsword:} is the most iconic weapon of knights, marking its users as people of status. Its cuts with the blade are very quick and effective against lightly armored opponents, in addition to having good reach and being usable against several opponents at once. The blade can also be used defensively to block. The longsword also has a sharp tip, which provides an attack with more penetration than the blade's cut. Another option is to thrust while holding the blade, reducing the reach but increasing penetration and allowing for attacks during grappling. Its tip is a precise attack, making it good for hitting vulnerable points. Furthermore, the sword also has a pommel, which can be used as an alternative for blunt damage. Longswords truly shine in the hands of more skilled characters.

\textbf{Polearm:} is the melee weapon with the greatest reach. Being a weapon capable of delivering braced attacks, it is excellent for keeping opponents at bay and discouraging charges. They have a shaft of approximately 3m in length, making them very inconvenient in enclosed spaces.

\textbf{Halberd:} is a long-shafted weapon with a tip composed of an axe head, a spear point, and a hook. The axe head delivers devastating attacks against objects of low rigidity. The spear point allows the halberd to be used as a short spear with two hands. The hook is used to pull opponents, knocking them to the ground or disarming them. 

\textbf{Warhammer:} is the most reliable melee solution against heavily armored opponents. It is like a halberd, but the axe head is replaced with a heavy hammer. Attacking with the hammer consumes a lot of energy and time, but if it hits, a heavy blow will knock down most opponents. This weapon is best used in the hands of a strong character.

\textbf{Greatsword:} is a larger and more robust version of a longsword. Although the weapon is quite heavy, it is balanced so that its attack can still be delivered with agility. The blade is very durable, making the weapon ideal for destroying other weapons that attempt to block it. Its sweeping attack with enhanced damage and long reach makes it excellent for dealing with many opponents at once. This sword is too large to be used in grappling combat, but it is still capable of thrusting at vulnerable points.

\end{multicols}
\begin{table}[H]
\scriptsize 
\centering
\renewcommand{\arraystretch}{1.4}
\hspace*{-1cm}
\begin{tabular}{ | c c c c c c c c c |}
\hline
\multicolumn{9}{|c|}{ \textbf{ \Large Two-Handed Weapons} } \\
\hline
Weapon & RES & Energy & SHP & hand & range & properties & price & size \\ 
\hline
\textbf{Longsword} & 20 & 12 & 1.5x & two & long I & sweep, DEF, heavy I, bladed & 12GP & large\\
% & 40/x2 & STR/1x & 3 & two & short & heavy I && \\
& 20 & 8 & 2x & one & short & piercing, penetrating, precise && \\
& 20 & 8 & 2x & two & long I & piercing && \\[5pt]
\hline
\textbf{Pike} & 20 & 8 & 2x & two & long II & braced, piercing, heavy I & 6 GP & large\\ 
& 20 & 0 & 1 & two & long I & shaft & & \\ [5pt] 
\hline
\textbf{Halberd} & 20 & 8 & 2x & two & long I & braced, piercing, heavy I & 12GP & large\\
& 30 & STR & 1.5x & two & long I & hook & & \\
& 20 & 0 & 0 & two & short & DEF, shaft, grapple & & \\
& 30 & 12 & 1.5x & two & long I & heavy I-III, bladed && \\ [5pt]
\hline
\textbf{Warhammer} & 20 & 8 & 2x & two & long I & braced, piercing, heavy I & 12GP & large\\
& 30 & STR & 1.5x & two & long I & hook && \\
& 20 & 0 & 0 & two & short & DEF, shaft, grapple & & \\
& 100 & 12 & 1x & two & long I & heavy I-III, smash && \\ [5pt]
\hline
\textbf{Bardiche} & 30 & 12 & 1.5X & two & long I & heavy III, sweep, bladed & 10 GP & large \\
\textbf & 20 & 0 & 0 & two & short & DEF, shaft, grapple & & \\ [5pt]
\hline
\textbf{Greatsword} & 30 & 12 & 1.5x & two & long I & sweep, DEF, heavy I-II, bladed & 18 GP & large\\
% & 100/x2 & STR/0 & 3 & two & short & heavy I && \\
& 20 & 8 & 2x & two & long I & piercing, heavy I, braced && \\[5pt]
\hline
\end{tabular}
\end{table}

\begin{multicols}{2}

\subsection{Ranged Weapons}

\textbf{Net:} it is a good weapon for ambushes or hunts. When thrown against groups of opponents, it can provide a decisive advantage by entangling several of them. Its preparation time is very long, making it impractical to throw multiple times in the same combat.

\textbf{Grappling Hook and Rope:} it is a weapon that, when it hits, allows you to pull a target. It is also a weapon that requires reload time, making multiple shots difficult. The grappling hook and rope can also be used as a climbing tool.

\textbf{Grenade:} it is a throwable explosive. The grenade can be magically activated in different ways to produce various effects.

\textbf{Short Bow:} it is the lightest bow. The strength required to draw it is low, allowing for faster shooting techniques. Short bows have a shorter range and deal less damage than other bows, making them poor choices against resilient opponents. Short bows can use broadhead arrows for more damage or bodkin arrows for better penetration.

\textbf{Long Bow:} it is the most common bow on the battlefield. It has enough power to penetrate medium armor when it hits directly. Drawing a bow like this requires above-average muscle strength.

\textbf{Heavy Bow:} this type of bow is so heavy that only the heaviest armors can stop its arrows. The strength required to draw such a bow is so great that very few humanoids are capable of using it. This weapon can use larger arrows, intended to take down large animals and heavily armored soldiers.

\textbf{Crossbow:} the crossbow is a weapon that can be fired quickly but requires reload time. The power of a crossbow shot is equivalent to that of a long bow.

\textbf{Arbalest:} the arbalest is an oversized crossbow that needs a lever-based reloading system to be drawn by an average person. The power of its shots is comparable to that of a heavy bow.

\end{multicols}
\begin{table}[H]
\scriptsize 
\centering
\renewcommand{\arraystretch}{1.4}
\hspace*{-1cm}
\begin{tabular}{ | c c c c c c c c c |}
\hline
\multicolumn{9}{|c|}{ \textbf{ \Large Ranged Weapons} } \\
\hline
Weapon & RES & Energy & AP & hands & range & properties & price & size \\
\hline
\textbf{Net} & 15 & grapple & 4+10 & two & 9m, 2m rad & grab, reload, throw & 3 GP & medium \\ [5pt]
\hline
\textbf{Grappling Hook and Rope} & 10 & 6 & 4+4 & two & 2xSTR & throw, hook & 2 SP & medium \\ [5pt]
\hline
\textbf{Grenade} & 20 & X & 3 & one & 2xSTR & throw & 2 GP & small \\ 
\hline
\textbf{Short Bow*} & 10 & 6 & 4 & two & 100m & piercing, fast & 5 GP & medium\\
\hline
\textbf{Long Bow*} & 10 & 8 & 5 & two & 150m & STR 13, piercing, fast & 8 GP & large\\
\hline
\textbf{Heavy Bow* } & 10 & 10 & 5 & two & 200m & STR 16, piercing, fast & 10 GP & large\\
\hline
\textbf{Crossbow} & 10 & 8 & 3+5 & two & 150m & reload, piercing, fast & 8 GP & medium \\
\hline
\textbf{Arbalest} & 10 & 10 & 3+6 & two & 200m & reload, piercing, fast & 12 GP & large \\
\hline
\end{tabular}
\end{table}
\begin{multicols}{2}

*Bows require training to be used effectively. Characters without proper training spend an extra 3 AP for each shot.

% ** Can use artillery arrows.

Arrows are small items that can be bundled into 10 units to make a small item. Bows and crossbows can use different types of arrows:

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item Bodkin: bodkin arrows have 3x SHP and is penetrating.
    \item Broadhead: has 2x SHP and is bladed.
    \item Explosive: gives the explosive effect to the arrow if ignited. Igniting is a standard action of prestidigitation.
\end{itemize}


\subsection{Shields}

\end{multicols}
\begin{center}
\begin{tabular}{||c c c c c c c||}
\hline
Name & Penalty & Deflection & RES & Insulation & Weight & Value\\ [1ex]
\hline \hline
Wooden Shield & -1 & +2 & 20 & 10 & 10 & 1 GP\\
Metal Shield & -2 & +2 &  30 & 10 & 25 & 15 GP\\
Wooden Tower* & -2 & +4 & 20 & 15 & 20 & 2 GP\\
\hline
\end{tabular}
\end{center}
Body shields are marked with * in the table and can provide total cover. 
% , but using them reduces running speed and long jump by 1m.

\begin{multicols}{2}

\section{Armors}

Armors have prot, which is used as damage reduction against impacts. The armor's TGH value is used to determine the breaking point of the armor. When there is no value for TGH, that means that the armor cannot be destroyed by an impact, which is the case for all non rigid armor. RES is both the armor breakage threshold and the damage reduction against CUT, since the armor must be destroyed to allow CUT damage to hit the wearer.

\textbf{Load penalties:} The penalties shown in the armor and shield tables derive from the additional weight of the armor and are reflected as a penalty in the AGI and STA attribute values. The penalties from multiple gear stack.

\textbf{Fiber armors} require the user to defend against most attacks actively, but they prevent the damage from being severe if hit. Cloth armors have very few penalties and a hardness of 2.

\textbf{Metallic armors} cause the character to sink 1m per turn while swimming, on top of whatever they get in their swimming test. Metal armors cannot be cut without the penetrating property.

\textbf{Rigid armors:} are layered to allow mobility. When RES is described with two numbers, the second one is the inner layer. Rigid armors receive and deal damage from weapons that hit them, thus having a RES value. Their protection against impact is greater, as the armor absorbs a larger portion.

\textbf{Gauntlets:} Gauntlets give -3 to prestidigitation, -3 to accuracy and climbing, but the armor protects the vulnerability of the hands. A pair of gauntlets costs 15 GP or 15\% of the armor's value.

\textbf{Closed helmet:} gives -3 to reflexes, -3 to cunning, and -3 to swimming, but armor bypass is impossible in the head. A closed helmet costs 15 GP or 15\% of the armor's value.

\textbf{Gauntlets and closed helmet:} It is possible to remove the gloves and closed helmets, but removing each of these eliminates their bonuses.

\end{multicols}

\begin{center}
\begin{tabular}{||c c c c c c c c||}
\hline
Name & RES & Protection & Insulation & Deflection & Penalty & Price & Weight \\ [1ex]
\hline\hline
Fiber &&&&&&&\\
\hline
Skin & 0  & 0 & 0 & +3 & 0 & 0 & 0\\
Clothing & 3  & 0 & 2 & +3 & 0 &  1 SP & small\\
Fur coat & 4  & 2 & 5 & +5 & 0 & 0 & medium\\
Thick Hide & 6  & 4 & 7 & +6 & 0 & 0 & large\\
Gambeson & 8 & 3 &  6 & +5 & 0 & 5 GP & medium\\
Padded Armor & 12 & 5  & 8  & +5 & -1 & 2 GP & large\\
\hline
Metallic &&&&&&&\\
\hline
Chain Shirt & 10  & 3 & 5 & +5 & 0 & 20 GP & medium\\ 
Hauberk & 12  & 4 & 6 & +5 & -1  & 30 GP & medium\\
Brigandine & 15/8 & 6 & 8 & +6 & -2 &  50 GP & large\\
\hline
Rigid and Metallic &&&&&&&\\
\hline
Half Armor & 20/10 & 10 & 7  & +5 & -2 & 60 GP & large\\
Full Armor & 20/10  & 10 & 8  & +7 & -3 & 100 GP & large\\
Reinforced Armor & 25/10 & 12 & 10  & +8 & -4 & 150 GP & large\\
\hline
\end{tabular}
\end{center}
\begin{multicols}{2}


\textbf{Common clothing:} is made of cotton or linen. It protects against very light injuries, such as bumping a shoulder against a rock or being scratched by a cat. Common people wear clothes of this type.

\textbf{Padded Armor:} is a garment with many layers made of cotton or linen, accompanied by boots and a metal helmet. It is very thick, which makes it heavy and reduces the user's mobility. This armor surprisingly offers good protection, but it is not made of a material designed to withstand many cuts. This is the type of armor a peasant would wear if called to fight and had to find something at the last minute. It is a very cheap item relative to the value of other armors.

A high-quality padded armor would be much more expensive but would improve resistance to cuts.

\textbf{Greatcoat:} is made from materials designed to protect the user in battle. It can be made of special leather or some durable plant fiber. Its great advantage is the mobility it provides, combined with reasonable protection against medium piercing weapons. If the user is hit directly by blades, the material will give way, but in the case of grazes, the damage should be completely mitigated. There may be variations of the armor with greater rigidity or less weight.

\textbf{Chain Shirt: } is the only metallic armor among the light ones. Its value is somewhat higher than the others because the difficulty of production is greater. The chain shirt is worn over a greatcoat.

\textbf{Hauberk:} the first and lightest of the medium armors. It is made of thicker rings than the chain shirt and provides better coverage for the legs and neck. As it is still a fairly flexible armor, it allows heavy impacts to injure its user, but it also offers few penalties to mobility.

\textbf{Brigandine:} is made of rectangular metal plates sewn inside a fabric vest. Since the metal plates can slide over each other, the cost to mobility is not so high. It can be complemented with a closed helmet and gauntlets during a major battle, but it can also be used without the gauntlets to allow free use of ranged combat weapons. Most professional soldiers use this type of armor in their daily lives, as it has a high cost but is not prohibitive, and it is also not so heavy that activities like climbing hills or trees become impossible.

\textbf{Half Armor: } is a lighter version of full armor. Several of the more restrictive pieces are removed so that the armor remains medium. As it retains the steel breastplate, it is very good against projectiles. Only those who find a gap in the armor will be able to injure the wearer. This type of armor is often chosen by archers in open field battles.

\textbf{Full Armor: } is the traditional armor of the medieval knight. Those who wear such armor want to survive a rain of steel or expend all their energy in attacking while letting the armor defend them. It comes standard with gauntlets and a closed helmet, providing superior coverage and preventing almost any attack from hitting a vulnerable point. Its thick steel plates offer high resistance against all types of damage, and its rigidity is so high that it is common for more fragile weapons, like swords, to break upon impact. Adventuring in this armor is extremely tiring. There is an even more exaggerated version of this armor, which is the reinforced one.

\subsection{Breakage and Repair of Equipment}

The equipment breakage function is optional.

\textbf{Equipment Breakage:} Objects can be destroyed by any kind of damage. The object's RES represents their resistance to both cutting and blunt damage. Destroying an object by burning is a special case that depends on the material. The order of magnitude of burning damage is much higher than what would be lethal to a creature, so it is not necessary to have a specific value for it.

Objects can deal and take damage whenever they collide with something else, whether the collision is caused by themselves or by the other object. Cutting damage is dealt exclusively by the attacking object to its target, while impact is dealt to both.

\textbf{Breakage throuh physical damage:} compare blunt and cutting damage from the attacking object to the RES of the receiving one. Whenever a piece of equipment can be cut and receives cutting or blunt damage >= RES, the equipment has a 1 in 6 chance to break. If it receives both >= RES, it has 50\% chance. If any single damage exceeds RES twice, the item breaks instantly.

\textbf{Breakage and Hardness:} Collisions involving materials of hardness 1 and 2 do not pose a risk of breakage to other materials.

\textbf{Weapon Breakage:} A broken weapon can no longer be used and costs 50\% of its original value to be repaired. Weapons that have a shaft require the replacement of the shaft.

\textbf{Armor Breakage:} When armor breaks, it is considered pitted. Repairing armor costs 20\% of the armor's value.

\textbf{Armorer:} Knowledge of armoring and the appropriate equipment are necessary to repair weapons and armor. Repairing metal objects requires a forge.

\textbf{Sharpening:} Is an action that takes one exploration turn and increases the weapon's SHP by 25\%. On critical hits, it increases by 50\%. Requires a sharpening stone. This does not work on precise weapons, because they are already considered sharp.

% \subsection{Materials}

% The table is a reference for the hardness of materials. It is used to define what can be cut by what. A higher hardness can cut a lower hardness. 

% \begin{center}
% \label{materiais}
% \begin{tabular}{||c c ||}
% \hline
% Name & Hardness \\ [1ex]
% \hline\hline
% \hline
% Water & 0 \\
% Butter & 1 \\
% Flesh & 2\\
% Linen & 3 \\
% Silk & 3\\
% Chitin & 4-6  \\
% Horn & 5  \\
% Wood &  5-6  \\
% Bone & 6\\
% Stone &  6-9  \\
% Steel &  8-10 \\
% Diamond & 11 \\
% \hline
% \end{tabular}
% \end{center}

\section{Characteristics and Conditions}


\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item Condition: Sharp - Weapon has +0.25x SHP until the end of the combat. It can be sharpened to a higher degree, increasing by an additional 25\%.
    \item Condition: Dull - Weapon loses 50\% of its SHP.
    % \item Characteristic: Hard: reduces RES by 25\%, but increases SHP of bladed weapons by 0.5x when sharpened.
    % \item Condition/Characteristic: Pitted - Armor has its coverage reduced by 3.
    \item Condition/Characteristic: Misfitted - Armor has its penalty increased by 1.
\end{itemize}

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    % \item Characteristic: Layered - Armor bypass strikes secondary RES instead of body.
    \item Characteristic: Reflective: Reduces damage from radiant-type attacks by 50\% before reaching insulation.
    \item Characteristic: Insulating: Increases coverage against electric-type damage attacks by +5.
    \item Characteristic: Inert: Doubles the armor's insulation against corrosive-type damage attacks.
    \item Characteristic: Sealed: Has infinite coverage against corrosive-type damage attacks.
    % \item Characteristic: Liquid: Immune to damage from slashing and bludgeoning-type attacks.
    \item Characteristic: Low-quality weapon - Weapon loses 25\% of its SHP and 25\% RES. -50\% cost.
    \item Characteristic: Low-quality armor - RES  reduced by 25\%. -50\% cost.
    \item Characteristic: High-quality weapon - Weapon has +25\% RES. +200\% cost.
    \item Characteristic: High-quality armor - RES increased by 25\%. +200\% cost.
    \item Characteristic: Semi-automatic - Has multiple shots per reload.
    \item Characteristic: Air filter - Closed helmet that grants immunity to toxic gas effects.
    \item Characteristic: Magnetic - Allows the use of magnetic abilities with the equipment.
    \item Characteristic: Living steel - Can be magically sharpened.
\end{itemize}


\section{Containers and Load}

Items can be categorized as small, medium, large, and cargo. 

\begin{itemize}
    \item Small items are the kind of thing you can hold in your hand. Tools, small weapons, coins, flasks and similar items are small. They tend to stay under 0.5 kg in weight and 0.5 liters in volume. It is possible to have a bundle of very small items that together are considered a small item, such as a bundle of arrows or a bag of coins.
    \item Medium items are normally cumbersome. They include one-handed weapons, folded tents, books, large bottles, etc. They tend to stay under 2 kg in weight and 2 liters in volume.
    \item Large items are at the limit of what can be carried. They include two-handed weapons, armors, boxes, tinder. They tend to stay under 10 kg in weight and 10 liters in volume.
    \item Cargo items are oversized items that require special handling and transportation, such as a horse or a cart. Their actual size is counted as weight and volume.
\end{itemize}

\textbf{Containers and burden:} Travelers need to have containers to carry equipment that they are not wearing or wielding. The containers impose a level of burden to be carried, determined by their size. They do no need to have anything inside to inflict the penalty. A character can be in one of four levels of burden: light, medium, heavy, and limping.

\begin{itemize}
\item \textbf{Light}: the burden has no effect.
\item \textbf{Medium}: -1 for burden penalty.
\item \textbf{Heavy}: -2 for burden penalties.
\item \textbf{Over}: -3 for burden penalties. Character is considered lame.
\end{itemize}

\textbf{Slot size and stacking:}Each container has a number of slots, which can each accomodate an item or stack of items. Slots are of one of three different kinds: small, medium, and large. A small slot can hold a small item. A medium slot can hold a medium item or a stack of 5 small ones. A large slot can hold a large item, a stack of 5 medium items, or a stack of 25 small items. Only identical items can be stacked together.

\begin{center}
\begin{tabular}{||c c c c c||}
\hline
Name & Slots & type & burden & cost \\ [1ex]
\hline \hline
Belt* & 4 & light & light & 1 GP  \\
Backpack & 12 & medium & medium & 5 SP  \\
Large Backpack** & 12 & large & heavy & 1 GP  \\
Sled & 200 & Cargo & over & 2 SP  \\
One-Horse Cart & 500 & Cargo & over & 30 GP  \\
Two-Horse Cart & 1000 & Cargo & over & 60 GP  \\
\hline
\end{tabular}
\end{center}
*items stored in the belt can be used with a standard action of prestidigitation
** STR 15 or size 4 reduces penalty to medium

A character can use only one backpack and one belt at a time.

\section{Wealth and Money}

Wealth is divided into two categories: currency and assets.

Copper, silver, and gold are everyday currencies. Ten coppers are worth the same as one silver, and 10 silvers are worth the same as 1 gold. 1 coin weighs 10g, which means 50 coins weigh 500g, equivalent to 1 small item.

Assets represent land, rare items, slaves, investments, a business, or another form of wealth that is owned but has a fluctuating value compared to regular currency.

\subsection{Salaries and Cost of Living}

The values of salaries and cost of living are described here are a guideline designed to ground the cost of things. 

\begin{center}
\begin{tabular}{||c c c c||}
\hline
Job & apprentice & specialist & master \\ [1ex]
\hline \hline
Worker & 1SP/day & - & - \\
Infantry & 2SP/day & 4SP/day & - \\
Archer & 2SP/day & 6SP/day & 12SP/day \\
Knight & - & 8SP/day & 3 GP/day \\
Armorer & 2SP/day & 5SP/day & 2 GP per day \\
Carpenter & 2SP/day & 3SP/day & 5SP per day \\
Mason & 2SP/day & 3SP/day & 5SP per day \\
% Priest & 2SP/day & 4SP/day & 8s per day \\
Doctor & 3SP/day & 1 GP/day & 3 GP per day \\
\hline
\end{tabular}
\end{center}

Any of these can be hired, as long as they accept the job.


Lifestyle:
1SP per day


\subsection{Magical Items}

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item Alchemical Torch - Allows the use of some light spells - Weight 2 - 12 GP.
    \item Superior Alchemical Torch - Allows the use of some light spells and automatically increases its size once - Weight 2 - 50 GP.
    \item Alchemical Prism - Allows the channeling of light - 150 GP.
    \item Electrite - Allows the transformation of magic into electricity, dealing 10 base electric damage. - Weight 2 - 30 GP.
    \item Superior Electrite - Allows the transformation of magic into electricity, dealing 15 base electric damage. - Weight 2 - 400 GP - carries 20 charges.
    \item Magnetite - Allows the transformation of magic into magnetism. - 10 GP.
    \item Fire Stone - Allows the production of fire from it, dealing 20 base burn damage. - 50 GP - carries 20 charges.
    \item Flame Thrower - A staff mounted with a fire stone to apply pressure to the flames, deals 20 base burn damage. - 80 GP. 
    \item Venomous Frog Corpse - Produces anticoagulant poison. - 5 GP - Weight 1.
    \item Soul Stone - Used to preserve souls - 200 GP.
    \item Condenser - Allows the extraction of water from the air. - Weight 2 - 5 GP.    
    \item Smoke Grenade - An item that can be thrown or attached to a projectile and produces smoke. 1 GP. Weight 1 for 5 units. 
    \item Explosive - Used for fire abilities. When ignited, it deals 5 fire damage with PEN 5 per turn of contact. 5 SP for 1 charge. 5 charges weigh 1. One charge can be affixed to an arrow.
    % \item Sharpening Sheath - Reduces the cost of the Sharpen spell on a weapon within the sheath to 1 AP - Weight 2 - 50 GP.
\end{itemize}

\subsection{Potions and Poisons}

Potions require a standard action to be consumed. Taking multiple potions does not amplify the effect. The contents of a potion on the list are relative to a creature of size 3. A creature of a different size will consume an amount equivalent to 1x Volume Modifier.

Potions add an effect to the character for a certain duration. The effects of potions are not cumulative, but they can be refreshed by consuming another potion of the same type. The effects of potions are not amplified by consuming multiple potions of the same type. Potions do not add hunger or thirst when healing.

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item Simple Healing: 4 cures over 1 exploration turn, 2 GP.
    \item Medium Healing: 8 cures over 1 exploration turn, 6 GP.
    \item Good Healing: 16 cures over 1 exploration turn, 16 GP.
    % \item Constitution: Increases constitution by +1 for 1 day, 4 GP.
    % \item Power: Increases power by +2 for 1 day, 4 GP.
\end{itemize}

\textbf{Poisons} apply an effect to the target. The magnitude of the effect is determined by the potency of the poison, the amount of poison applied and the target's resistance to the poison, represented by Health. The amount of poison applied is determined by the method of application (some methods may increase the amount of poison applied) and the seriousness of the injury.

\textbf{Dosage:} Tier 1+ wounds from PEN apply the full effect of the poison. Tier 0 decrease DL by 3. Inhalation applies the normal DL, independent of creature size.

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item Anticoagulant Poison: DL 5 health when applied to get 1 point in hemorrhage, 10 GP for a vial with 10 applications and weight 1.
    % \item Paralytic Poison: DC 10 health to avoid being paralyzed and lose 100/50/25/0 \% of maximum AP for 1 hour or until receiving an antidote.
\end{itemize}


\subsection{Utility and Survival}

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item Small Tent: Allows you to set up camp without skill. Has space for 2 people and reduces exposure by 3 - 1 GP - medium
    \item Large Tent: Allows you to set up camp without skill. Has space for 4 people and reduces exposure by 3 - 2 GP - large
    \item Flint: Grants +5 on the test to light a campfire. 2 PP - small
    \item Firewood: Wood ready to burn for 8 hours. Grants +5 on the test to light a campfire. - medium
    \item Rope: Useful item. 1 PP per meter. Medium
    \item Canteen: Holds 2L of water. 1 PP. Small.
    \item Jug: Holds 5L of water.  1 PP. Medium.
    \item Shovel: Used for digging. 1 PP. Large.
    \item Pot: Used to boil water and cook up to 10 portions, improving the preparation test by 2.  5 PP. Medium
    \item Vial: Holds 50ml of any liquid. 5 CP. small
    \item Alcohol - Used for fire skills. 1 PP per 250ml (5 charges). 
    \item Oil - oil. 1 PP per 250ml (5 charges). 
    \item Sharpening Stone - A sharpening stone. 2 PP. small 
    \item Fur Cloak - Improves resistance to cold wear by +3. 1 GP. medium
    
\end{itemize}
    
\subsection{Drugs}

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item Popaine:  - 5 PP - Causes the character to become confused or, in high doses, incapacitated.
    \item Alcohol: Causes the character to become intoxicated. 
    
\end{itemize}


\end{multicols}
