% !TeX root = main.tex
\chapter{Introduction}

\begin{multicols}{2}


\textbf{How the book should be used:} 
- I only added what was necessary for tactics and operation of core game mechanics. Many things are blueprints for personalization.

\textbf{Rules vs Guides:} Rules are necessary to give players agency. Guides are used to instruct about how the game was designed to be played. They are not necessary for the game to work.

\textbf{Player skill vs character skill:}

\textbf{Tests and the role of dice:} Skills are the mechanism used to perform tests. All tests use one of the 20 skills.

\section{Types of games}

\textbf{Historical:} The magic system is optional and can be ignored. This game is good for any time period from antiquity to pike and shot. It is not really designed for modern warfare, since much of the mechanics would become obsolete and unimportant in the presence of modern weaponry.

\textbf{Horror:} Stress mechanics. Characters lose control. Lighting. Time.

\textbf{War:} Tactical movement of troops. Technological race. Supply and logistics management. Disposable characters. Helpers. 

\textbf{Survival and adventure:} Exploration mechanics. Management of inventory, sleep, thirst, and hunger.

\textbf{Intrigue:} Social. Progress in relationships with NPCs. Social skills. Reputation.

\section{Types of combat}

\textbf{Deadly:} Ends when one team is neutralized.

\textbf{Defend:} Ends when time runs out and the target is protected.

\textbf{Defeat Target:} A specific target must be eliminated. One can flee after defeating the main target.

\textbf{Theft:} The objective is to acquire an object. 

\textbf{Escape:} The objective is to evade pursuers or reach a safe place.

\textbf{Stealth:} The objective is to go unnoticed.

\textbf{Escort/Kidnapping:} The objective is to transport a character.

\textbf{Social} Debate, negotiation, seduction, intimidation. 


\end{multicols}