\chapter{Sorcery}

\begin{multicols}{2}

\newcommand{\spell}[9]{
    \textbf{\large{#1}} \\
     \textbf{Cost:} #2 \\
     \textbf{Talent:} #3 \\
     \textbf{Requirements:} #4 \\
     \textbf{TN:} #5  \\
     \textbf{Range, Area:} #6 \\
     \textbf{Spell Type:} #7 \\
     \textbf{Duration:} #8 \\
     \textbf{Effect:} #9\\
}

\section{Sorcery basics} 

Sorcery is the ability to affect things outside of self using only the mind. Sorcery tests can be used on any magical artifact that can be affected. 

Spells are just useful abilities that a character can train and improve at for repeated usage. They do not represent everything that can be done with sorcery.

\subsection{Learning spells}

Spells can be learned in three different ways: intuitively, as a wizard and as a cleric.

\textbf{Intuitively:} A specific spell can be learned through practice. Doing it this way costs 6 XP, and the talent used to learn is the spellcasting proficiency level. It is not possible to learn spells with knowledge requirements above 3 this way.

The skill to perform spells this way is pure spellcasting proficiency. Spells learned intuitively are allowed to be cast safely. The spell can be learned again for 2 XP to decrease its difficulty by 1 and then again by 3XP to reduce it by 2 and so on increasing the requirement by 1 to reduce it further by 1 point. This can be done to reduce difficulty up to 10 points.

\textbf{Wizard:} A wizard is a scientist of magic. They learn their spells through books and experimentation. The knowledges related to magic are Alchemy, Animancy and Biomancy. The \textbf{Knowledge level + Sorcery} is used as a skill to perform spells.

Wizards can attempt to interact with anything that has magical properties, without prior learning. This works like a regular sorcery test, but must be done in a risky manner. For every knowledge point missing compared to the object's requirement, increase difficulty by 5.

The wizard can learn spells from grimoires or through experimentation within an exploration turn, given that the required materials are present. Make a sorcery test and learn on a hit or crit. It has its difficulty increased by 5 for every knowledge point missing. 

\textbf{Cleric:} can cast every spell in the deity's list, as long as it has the required devotion level. The \textbf{Devotion level + Sorcery} is used as a skill to perform divine spells.

Clerics can also attempt to cast miracles. When they do that, they use \textbf{2x Devotion} as a skill, and can try to cast any spell, regardless of devotion level, but get a -3 penalty on the test per devotion level under the spell's talent requirement and must perform them in a risky manner. It is impossible to cast spells this way in places outside the influence area of the deity. In a temple or holy place, the effect is opposite, and the difficulty to cast is reduced. 

\subsection{Casting spells}

Casting a spell is a skill test against a spell TN. Any points above the TN are converted into SOPs, which can be used to improve spells, as explained in the spell modifications. 

Casting a spell costs AP and STA during combat or exhaustion during exploration. To cast any spell, make a sorcery test against the \textbf{Spell Casting TN} to determine if the spell was successfully cast or not. In case of failure, it fails and the AP and STA are lost. 

\textbf{Concentration:} All sorcery tests require concentration. No other action or reaction can be performed while concentrating. Being interrupted causes the loss of concentration. Sometimes, a character's AP may not be enough to cover the action's requirements. In that case, they can maintain concentration and continue acting in the next turn. Losing concentration causes all AP and STA spent to be lost. 

\textbf{Sustained Spells:} are actions that can be performed continuously while concentration is maintained. Their effect ends when the caster stops concentrating on it. For these spells, the sorcery skill test needs to be redone periodically, as indicated in the spell. A failure in the sorcery test interrupts the sustained magic.

\textbf{Charged:} can be charged into something and released when a trigger is activated. If the spell is not used within the duration of the charge, it dissipates. 

\textbf{Delayed Spells:} Any spell can be delayed, meaning that the casting can be completed, but the release is postponed. The character must concentrate on the spell until the release, which can be triggered by a previously prepared action. If a trigger does not occur, remake the sorcery test every turn to keep the spell ready.

\textbf{Requirements:} Some spells may even be impossible to cast if the minimum conditions are not met. Making it rain in the desert is impossible if there is no water in the sky.

\textbf{Stealth and Sorcery:} Spells are imperceptible through normal senses before they cause their effect. Once their effect occurs, they can be identified with a test of the skill used to cast the spell.

\subsection{Types of Spells and Modifications}

Enhancements are modifications that casters can make to their spells to make them more powerful or better suited to their needs. For every point over the spell's TN, get SOPs that can be used to improve spells.

\textbf{Accelerate Spell:} Spend 3 SOPs to Accelerate a spell for 1 AP. It can be used multiple times to reduce the sorcery time to up to half the original value. It is possible to accelerate spells to avoid opportunity attacks.\\

\textbf{Extend Spell:} Spend 2 SOPs to increase the range of a spell according to the MM.

\textbf{Enhance Spell:} Spend 4 SOPs to improve spells in a special way described in the spell.  \\

\textbf{Amplify Spell:} Spend 4 SOPs to amplify spells. It can be used when the spell has its effects multiplied by DM, SM, MM, and VM. The number of times it can be increased depends on environmental conditions and the material used. The amount of material or charges consumed is multiplied by the VM. Increase Spell allows it to apply to larger versions of the items. \\

\textbf{Extend Charge Duration:} Spend 2 SOPs to increase the charge duration of charged spells by +50\%, then +100\%, then +200\%, +400\%, and so on in progression. \\

\textbf{Effortless Spell:} Spend 4 SOPs to rest while casting a spell. The spell must spend the higher between the rest's AP cost and the spell's AP cost. If used outside combat, reduces the exhaustion cost of a spell by half.  \\

The types of spells are:

\begin{itemize}[noitemsep]
\item Environmental: The spell alters the environment. They have customized enhancements.
\item Effect: Spells that directly affect the target and have various effects. They have customized enhancements.
\item Psychic: Spells that directly affect the target and have psychic effects. They use the Charisma proficiency as the Difficulty Level (DL). Enhancing increases the DL by 2.
\item Projectile: Spells that function as projectiles, using the shooting skill as the DL.
\item Touch: Spells that function as strikes, using the striking skill as the DL. They do not trigger opportunity attacks.
\end{itemize}

\section{Alchemy}

The art of giving life to objects and manipulating life in them. The power and efficacy of any alchemical spell is based on the gear used to perform it.


\spell{Detonate Explosive}{1 charge, 2 AP + 1 STA}{Alchemy 2}{ explosive}{5}{15m}{Effect}{}{ Detonates an alchemical explosive that is already loaded, causing it to deal its damage and effects. }

\spell{Smoke Explosive}{1 grenade, 3 AP + 1 STA}{Alchemy 3}{ explosive}{10}{Explosion with a 5m xMM radius}{Effect}{loaded, 5 rounds}{ Loads explosives with the effect of producing smoke when activated, upon contact with fire or on impact. The smoke renders anyone within the fog invisible to the naked eye and obscures visibility to a radius of 1m. }

\spell{Shock Explosive}{3 charges, 4 AP + 1 STA}{Alchemy 3}{ explosive}{10}{Explosion with a 3m xMM radius}{Effect}{loaded, 5 rounds}{ Loads explosives with alchemical energy to explode when activated, upon contact with fire or on impact, dealing 20xDM of impact damage and 20xDM PEN as an explosion.}

\spell{Incendiary Explosive}{3 charges, 4 AP + 1 STA}{Alchemy 3}{explosive}{10}{Explosion with a 3m xMM radius}{Effect}{loaded, 5 rounds}{ Transforms the explosive into a sticky substance that ignites when activated, upon contact with fire or on impact, dealing 30xDM burn damage. }

% \spell{Extinguish Flames}{ 2 AP + 2 STA}{Alchemy 1}{ Alchemy 4}{8}{Touch}{Environmental}{instantaneous}{ Extinguishes the flames of any combustible material it can touch.}

% \spell{Start Fire}{ 1 charge, 3 AP + 3 STA}{Alchemy 1}{igneous stone}{5}{Touch}{Environmental}{instantaneous}{ Sets fire to any combustible material it can touch. }

\spell{Flamethrower}{6 AP + 1 STA, + 1 charge per activation}{Alchemy 3}{Flamethrower}{8}{Cone explosion of 4m xMM in length and 60 degrees in radius}{Projectile}{sustained}{ Creates flames from an igneous stone, dealing its base damage xDM in burn damage to all within a cone in front of the caster. The flamethrower can be sustained, dealing explosion damage to anyone entering the area. The damage is applied only once, based on the position that causes the most damage. }

\spell{Fire Breath}{2 AP + 1 STA}{Alchemy 2}{ Igneous Implant }{8}{ Long I }{Touch}{instantaneous}{ Allows the user to blow hot air from their mouth as if it were a long weapon. The attack deals burn damage equal to the base damage of the implant xDM. This attack is considered precise and a small weapon. }



\spell{Extract Water}{ritual, 1 Exhaustion}{ Alchemy 1 }{Condenser}{variable 0-15}{melee}{Ritual}{4 hours}{Extracts moisture from the air and directs it to a location, which can be a container. The amount of water extracted is 1L for matching the TN and another 1L for every point over it. Increase or decrease TN based on available humidity.}


% \spell{Toxic Gas}{6 AP + 3 STA}{ 3}{Medicine II}{14}{5m, radius of 3m}{Environmental}{3 turns, loaded}{Causes a corpse to release poisonous gas for the entire duration. Undead count as corpses. Anyone breathing within the area at the start of their turn makes a Health test against the spell's DC to avoid coughing. Loses 6/4/2/0 AP, depending on the test result.}

% \spell{Fire Storm}{5 AP + 6 STA}{ 4}{fuel}{16}{10m radius centered on the caster }{Environmental}{variable}{Ignites any combustible object in the area. It does not deal damage initially, but blinds and suffocates anyone inside and deals 5 burn damage with PEN 4 to anyone who starts their turn within it.}

% \spell{Caixão de Areia}{6AP + 8 STA}{ 4}{}{16}{directly from 30m}{Effect}{5 turns}{Attempts to cover a creature of size 4 or smaller with sand. The sand follows and grabs the target, paralyzing and suffocating for 1 turn if they fail or scratch their Strength or Evasion defense. The effect is reapplied each turn until the duration ends. Requires large amounts of sand.}

\spell{Repair Metal}{ Ritual 1 min }{ 3}{ living steel }{3}{touch}{Effect}{ 1 min }{ Repairs cracks in metals. Works like a cold weld, removing negative conditions from equipment. }

\spell{Metal Sharpening}{ 3 AP + 1 STA }{Alchemy 3}{living steel weapon}{5}{touch}{Effect}{charged}{ Performs a sharpening action on a weapon. Improving gives a critical hit on the sharpening.  }

% \spell{Galvanic Corrosion}{3 AP + 1STA or 1 exhaustion for 4kg}{Alchemy 2}{electrite}{5}{melee, target}{Touch}{charged, 1 turn}{ Rusts an object that is susceptible to rust. Leaves armors pockmarked and weapons smaller than size 5 without wire. It is necessary to hit an attack against an object for the effect to occur.}

\spell{Sustained Lightning}{4 AP + 1 STA, 3AP+1STA per activation, 1 charge per activation }{Alchemy 5}{electrite}{8}{9mxMM}{Projectile}{sustained}{The caster attempts to create a sustained electric arc between themselves and a target, causing the base damage of the electrite x2DM electric damage. Once the spell hits, it only ends if the target leaves the spell's range or touches the caster, in which case the arc short-circuits, causing damage to both and ending the spell.}

\spell{ Taser }{2 AP +1 charge}{Alchemy 1}{electrite}{5}{touch}{Touch}{charged, 1 turn}{ Charges an object of electrite with electricity that discharges on the first object it comes into contact with, adding the electric damage of the material to the attack. }


% \spell{Phantom Sound}{3AP + 2 STA}{ 1}{}{15}{30m}{Environmental}{instantaneous}{The wizard makes a sound propagate from a point at a distance they do not need to see, only pointing in the general direction. The target must make a perception test to try to immediately know it is an illusion; otherwise, they will react normally to the sound. The sound can be as loud as a scream.}

\spell{Recharge Stone}{1 charge per 10 minutes, 3 exhaustion per TE}{Alchemy 1}{stone}{5}{Effect}{}{charged}{ Recharges charges of a stone. Ateu devotees can cast this spell without Exhaustion cost.}

\spell{Darken}{3 AP+1 STA per turn }{Alchemy 3}{ Alchemical Prism}{8}{1m radius}{Environmental}{sustained}{Absorbs all light entering an area, making it difficult for those outside to see what is inside. Makes anyone inside invisible to those outside, who only see a large black circle.}

\spell{Lasers}{5 AP + 2 STA, 3 AP + 1 STA per activation}{ Alchemy 4}{ Alchemical Prism}{12}{200m xMM}{Projectile}{sustained}{Converges the rays of light from a large light source into a single point at a distance, causing 30xDM radiant damage to the target per activation. The laser is considered a projectile but does not receive any distance penalty up to 50m.}

% \spell{Invisibility}{3 AP +1 STA}{Alchemy 4}{ Alchemical Prism}{12}{self}{Environmental}{sustained}{Bends the rays of light around the caster during the duration to make them invisible.}



\spell{Lamp}{3AP or as a ritual 1 exhaustion per PV}{Alchemy 1}{ alchemical torch, }{5}{touch, 10m radius}{Environmental}{charged, 1 minute}{Creates a ball of light in a metallic object that you can touch. Otherwise, it makes the caster themselves glow, but costs double stamina and PV.}

\spell{Dazzling Flash}{4AP+1STA}{Alchemy 2}{ alchemical torch}{8}{self, 6m radius cone at 30 degrees}{Environmental}{instantaneous}{Produces a powerful flash that blinds anyone in the area. Victims must make a reflex test, at no AP cost. On a failure or graze, they are dazzled for 1 round. The flash is blocked by fog, dust clouds, and magical darkening.}

\spell{Revealing Light}{4AP or 1 exhaustion per TE }{Alchemy 2}{alchemical torch}{5}{touch, 10m radius}{Environmental}{charged, 1 minute}{A bluish light emanates, revealing all kinds of mysteries and invisible things. Transforms anything invisible within the area into something visible.}

% \spell{Radar}{4AP }{Alchemy 5}{alchemical torch}{10}{touch, 200m radius}{Environmental}{instantaneous}{ Can see through walls and thin barriers in a large radius for an instant. Can see invisible things. Can be improved multiple times by spending 1 STA for each and to increase the radius by 200m.}

\section{Animancy}

This is the art of projecting the soul into objects and creatures and injecting fragments of consciousness into them. 

% Animancy requires some for of weakness of the mind to be used. Create relation with morale system.

% Area effect magic

% Spirit binding and resonance

% Scouting for spirits and communicating

\spell{Anchor Soul}{Ritual}{Animancy 3}{Soul stone}{8}{melee, target}{Effect}{3 days, sustained}{Anchors a creature's soul to a soul stone, preventing it from disappearing. This can be done with the souls of creatures that have been dead for up to 10 minutes. If the soul is still present, resurrection is possible, but the soul will extinguish in 3 days without a body.}

\spell{Spiritual Strike}{3 AP + 1 STA as a projectile}{Animancy 2}{ }{5}{15m}{Projectile}{}{Launches part of your soul violently at an enemy, interrupting them if the caster's animancy + 3 SM is greater than the target's will + 3 SM. If it is a critical hit, gain +3 animancy. The projectile is invisible to anyone who cannot detect spells.}

\spell{Spiritual Walk}{5 AP + 2 STA, +2 STA per turn}{Animancy 3}{Synesthesia I}{8}{100m}{Effect}{sustained}{Exits the body in spirit form, leaving it unconscious. This form travels in any direction with a movement speed of 6m. It is invisible and immune to any type of damage and can cast spells normally, but the character spends 2 STA per turn while out of the body, or 1 STA per turn if perfected. The spirit's line of sight is equal to the basic version of the Synesthesia ability.}

\spell{Ghost}{5 AP + x STA, 1 exhaustion}{Animancy 3}{Synesthesia I}{10}{self}{Effect}{instantaneous}{Can generate a ghost that behaves like the spiritual walk but has limited energy. The ghost receives x STA and disappears if STA reaches 0. All memories of the ghost are lost if it does not return to the original body. The ghost loses 2 STA per turn or 1 STA if the spell is perfected.}

\spell{Disintegrate Mind}{8 AP + 8 STA}{Animancy 4}{ }{10}{melee, target}{Effect}{}{Destroys a creature's mind permanently. Reduces SPI to 0 permanently and resets all convictions to 0. This can be done by clerics of Iskar even without having learned this ability, but the target's consent is required.}

% \spell{Chaotic Thorn}{2 AP + 2 STA}{Animancy 1}{}{5}{30m}{Effect}{instantaneous}{As a reaction after seeing an enemy casting a spell, fires a chaotic thorn that increases the difficulty of casting the spell by +3.}

\spell{Interference Matrix}{3 AP + 3 STA or ritual 1 hunger for 5 minutes}{Animancy 4}{}{10}{6m radius, 10m range}{Effect}{5 turns}{Increases the difficulty of casting any spell within the area by 3.}

\spell{Possession}{4 AP + 4 STA}{Animancy 4}{ }{10}{melee}{Effect}{sustained}{Attempts to enter and possess another body during the spiritual walk or as a ghost. This is only possible if the body is confused, unconscious, or if the target allows the possession. The target is Dominated for the duration. Each time the possessed makes a will test, the possessor must spend 1 STA to resist. Possession ends when the possessor's STA reaches 0 or the possessed succeeds on a will test.}

\spell{Sleep}{3 AP }{Animancy 2}{}{5}{15m}{Psychic}{instantaneous}{ Causes someone who is not afraid or angry to fall asleep or to be asleep in combat. Targets fall asleep on a failure or a graze on a will test against the charm.}

\spell{Alter Memories}{5 AP}{Animancy 4}{}{10}{melee, target}{Effect}{ }{ Touches the mind of a person and sees the memories of the things the person is talking about. Anything said by the caster to the person will be remembered as truth. The person can accept the ritual or resist it with will. In this case, a success or critical modifies the memory successfully. If the person accepts the ritual, a cleric of Rivka automatically succeeds in the ritual. }

\spell{Psychoactive Assistance}{3 AP + 1 STA}{Animancy 4}{Central Intelligence}{10}{15m}{Effect}{instantaneous}{ Focuses on a creature with whom they share a telepathic link. The caster can make an insight or will test to resist or remove a mental status effect on behalf of anyone connected by a telepathic link. }

\spell{Frighten}{3 AP }{Animancy 1}{}{5}{15m, detectable target}{Psychic}{instantaneous}{ A spiritual attack that causes someone to become afraid if they fail or graze on a will test against the Stress.}

\spell{Calm}{3 AP }{Animancy 1}{}{5}{15m, detectable target}{Psychic}{instantaneous}{ A spiritual attack that removes fear or rage if they fail or graze on a will test against the Enchantment.}

\spell{Command}{6 AP +1STA}{Animancy 5}{Suggestion}{12}{15m}{Psychic}{up to unconscious}{ Gives a simple command to a confused target and leaves them Dominated if they graze or fail on a will test. They remain Dominated until they complete the task, become unconscious, or succeed on a will test to break free from the domination. It is not possible to change the task without using Command again, but it is possible to specify it. }

\spell{Pain}{3 AP }{Animancy 3, Biomancy 1}{}{8}{15m}{Psychic}{instantaneous}{ Causes a great sensation of pain, which paralyzes a target for 6 AP on a failure or 4 AP on a graze on a will test against the stress.}

\spell{Charm}{3 AP }{Animancy 1}{}{5}{15m, detectable target}{Psychic}{instantaneous}{ A spiritual attack that leaves the target enchanted if they fail or graze on a will test against the Enchantment.}

\spell{Visual Interference}{3 AP }{Animancy 3, Biomancy 1}{}{8}{15m, detectable target}{Psychic}{1 turn}{ Causes blindness in a target that fails or grazes on a will test. }

\spell{Telepathic Link}{3 AP}{Animancy 3}{}{5}{20m}{Effect}{10 minutes}{ Creates a telepathic link with a nearby creature if they fail or graze on a will test or if they allow the link. The caster can speak directly to the mind of their target during the duration of the spell or until the maximum distance is exceeded, at which point the spell breaks. Each simultaneous link increases the difficulty of the spellcasting test by 2.}

\spell{Offend}{3 AP }{Animancy 1}{}{5}{15m, detectable target}{Psychic}{instantaneous}{ A spiritual attack that causes someone to become enraged if they fail or graze on a will test against Stress.}

\spell{Psychosis}{3 AP + 1 STA}{Animancy 4}{ }{10}{15m, detectable target}{Psychic}{ }{ Causes confusion in a target that fails or grazes on a will test against the Stress. }

\spell{Mental Palace}{5 minutes}{Animancy 4}{}{5}{target, melee}{Effect}{5 minutes, sustained}{ Helps someone remember something, gaining +2 on any knowledge test. Additionally, they can recall all the details of anything they can perceive during the duration of the spell.}

\spell{Mental Paralysis}{5 AP + 2 STA}{Animancy 5}{}{12}{15m}{Effect}{1 turn}{ Completely paralyzes the mind of a confused creature that grazes or fails on their will test against stress, rendering them immobile and possibly fallen.}

\spell{Suggestion}{3 AP }{Animancy 2}{}{5}{15m target that can hear and understand the suggestion}{Psychic}{}{ Tells a creature that can hear and understand the words and is confused, enraged, enchanted, or frightened to do something, causing the compulsion to obey. If they fail or graze on a will test against Enchantment, they will obey. This compulsion should not cost more than 6 AP on a failure, 4 AP on a graze. }

\spell{Rescue}{3 AP+ 1 STA}{Animancy 2}{}{5}{15m, detectable and enchanted target}{Psychic}{instantaneous}{ Compels an enchanted creature to protect its seducer for 8 AP on a failure in the will test against the Enchantment or for 6 AP on a graze. }

\spell{Torment}{8 AP + 3 STA}{Animancy 5}{}{12}{15m, direct target}{Effect}{until they wake up}{ Infects an unconscious target within range with a terrible and extremely realistic dream, which keeps the victim asleep until they can escape the dream by achieving a success or critical on a will test against the Stress. They can attempt to wake up every hour. The victim does not heal from this sleep and receives 2 exhaustion for each hour they remain asleep.}

\spell{Overwhelming Terror}{3 AP + 1 STA}{Animancy 3}{}{8}{10m}{Psychic}{}{ Forces all frightened enemies to make a Morale test and flee from the caster for 4 AP on a graze, 6 AP on a failure.}

\spell{Provocative Mockery}{3 AP +1STA}{Animancy 3}{}{8}{15m}{Psychic}{}{ Forces enraged and confused targets to attack the caster for 6/4 AP on a failure and graze, respectively, on a will test against the stress.}


% Biological

\section{Biomancy}

Manipulate drugs, cells, etc. Mutations, grafts.

% poison, decomposition, and undead

% healing

\spell{Healing Food}{5 minutes}{Biomancy 4}{}{12}{touch}{Effect}{sustained}{ Modifies a portion of food to behave like a potion, giving half of its nutritional value as healing instead of hunger. }

\spell{Immune modulation}{ 1 Exhaustion }{Biomancy 3}{}{8}{melee, target}{ Effect }{ 10 minutes }{ Attempts to manipulate the body of a living being to metabolize poison. Allows a redo of a health test to resist poison using the target's health +2, and an additional +2 for each enhancement. }

\spell{Defibrillate}{3 AP + 1 STA}{Biomancy 2}{}{5}{melee, target}{Effect}{instantaneous}{ Wakes an unconscious target.}

\spell{Reattach Limb}{2 Exhaustion, 1 hour}{Biomancy 5}{}{12}{melee, target}{Effect}{}{Reattaches a limb to a body, adding a serious wound that requires 20 cures to function perfectly again.} 

\spell{Healing Touch}{4 AP + 1 STA}{Biomancy 2}{}{5}{melee, target}{Effect}{instantaneous}{ Clots 1 point of bleeding in a target that can be touched. Enhancing increases the amount of heals by 1. }

\spell{Regenerate}{15 minutes/cure, 1 exhaustion/hour}{Biomancy 3}{Medicine 4}{10}{melee, target}{Effect}{sustained}{Actively manipulates the functions of a target's body to accelerate healing. Removes 1 IL or heals a wound by 1 for every 10 minutes.}

\spell{Revitalize}{2 Exhaustion}{Biomancy 5, Animancy 3}{ }{15}{melee, target}{Effect}{5 minutes, sustained}{ Performs a 5-minute ritual to revive a creature within reach of the body, as long as the body is in a condition to be alive and there is a sample of the soul nearby. The body is revived with all its wounds. }

\spell{Enhanced Sleep}{1 TE, 1 Exhaustion}{Biomancy 1}{}{8}{touch melee}{ Ritual}{4 hours}{Blesses the sleep of everyone in the area for 4 hours, allowing everyone except the caster to receive +2 health for resting and to complete it in 4h.}

\spell{Preserve Corpse}{10 minutes}{Biomancy 1}{}{5}{melee}{Effect}{1 day, charged}{Prevents decomposition from affecting a corpse or part of it for a full day.}



% \spell{Nervous Shutdown}{4AP + 4 STA}{}{16}{30m direct}{Effect}{1 turn}{Attempts to magically block nerve communication through the nerves of a target, paralyzing them if they fail a challenge test and stunning them for 8 AP if they graze and for 4 AP if they succeed.}

% \spell{Paralyzing Infection}{6AP + 6 STA}{12}{14}{30m, radius of 4.5m}{+1/2/4 DL}{3 turns}{Releases a paralyzing parasite into the air and directs it towards a group of enemies. It weakens those who fail or graze a Health test in the first turn, then stuns for 5 AP enemies who fail or graze a second turn, and in the next turn paralyzes all who fail or graze a third Health test for 3 turns.}

% \spell{Corpse Explosion}{4AP + 3 STA}{5}{20}{60m, radius of 4.5m}{+0.5/1/2x damage}{instantaneous}{Destroys a corpse, dealing damage equal to 2x POD up to half the total HP of the corpse as bludgeoning damage.} \

% \spell{Decomposition Gas}{2 AP, 3STA}{5}{20}{9m, explosion terrain radius of 6/9m}{+0.5x poison damage}{instantaneous}{Infects anyone in the area with a toxic magic, causing 4 + POD/2 poison damage per turn and blocking healing effects. Deals double damage to undead.} \ ability

% \spell{Necrofire}{3 AP + 6 STA}{}{14}{15m}{Effect}{instantaneous}{Ignites enemies with a special energy that consumes flesh, dealing spiritual damage equal to 25\% of their missing HP on a graze and 50\% on a failure. If cast on undead, it cannot be resisted and deals 2x POD burn damage.}

% \spell{Accelerate Fermentation}{2 Hunger}{ 1}{Entropy I}{15}{melee, target}{Ritual}{30 minutes}{ Greatly accelerates the fermentation process of sugary substances, creating 500ml of alcohol.}

\spell{Animate Dead}{6 AP + 2 STA or as a ritual 1 min}{Animancy 4, Biomancy 4}{ Fresh body }{10}{melee}{Effect}{10 minutes, sustained}{Animates a corpse with a fragment of its soul. It can only move and attack or grapple enemies. The caster can direct it to attack something, but selecting a new target costs 1 AP. If the target dies, the corpse will automatically seek another. Multiple animated corpses can be commanded at the same time. The corpse retains its strength, size, and resilience, but the rest of the attributes become 6. If out of combat, it is possible to attempt to speak with the undead. If the brain and vocal cords are intact, it can answer some questions.}

% manipulate trees

\spell{Tree House}{1 ET, 2 exhaustion}{Biomancy 4}{}{10}{target}{Effect}{sustained}{ Grows trees and shapes them to have enough hollow space inside to serve as a room. The effect of the spell is permanent, but it takes the duration of the spell to shape the trees. The internal space is equal to 6 cubic meters xMM.}

\spell{Accelerate Blooming}{2 Exhaustion, 1 ET}{Biomancy 2}{}{5}{melee, target}{Effect}{sustained}{ The plant will go through its seasonal stages very quickly, draining nutrients and water from the soil and bearing fruit if possible. }

\spell{Defoliate}{1 min}{Biomancy 2}{Pheromone Control}{8}{12m, radius of 6m}{Environmental}{}{ Causes all deciduous trees in the area to lose their leaves within 1 turn. The leaves are flammable. }

\spell{Insect Pheromone}{5 AP + 1 STA}{Biomancy 3}{Pheromone Control}{8}{ radius of 3m x MM of taret}{Environmental}{}{ Secretes pheromone, which can be stored. Causes nearby insects to swarm anyone that gets in contact with it and sting them. The effects depend on the insects that are present and how long they will take to arrive. In combat, it works if there are insects within 50m, they take 1 turn per 10m to arrive.}

\spell{Stimulate Secretion}{ 1 min }{Biomancy 2}{dead animal}{5}{touch}{Effect}{}{ Forces a dead venomous animal to secrete venom. The venom produced is sufficient for 1 application and the effect depends on the corpse used. }

% \spell{Healing Avatar}{5AP + 3 STA}{  5}{Ghost, Healing Touch}{12}{30m, target}{Effect}{3 turns}{Sends a generous amount of healing energy to a target, performing 8 AP in healing spells per turn with the target's STA, but without costing them AP. }


\spell{Clone}{ 1 ET, 4 exhaustion }{Biomancy 6}{living sample, preserved corpse }{15}{melee}{Effect}{sustained}{ Converts the genes of a part of the body of a preserved corpse to match those of another creature of the same species. A living sample of the model must be present. Converts 10kg of biological matter per TE.}



% \spell{Body Sculpting}{ 4 Hunger, 1 hour}{Biomancy 5}{Medicine 5}{12}{Melee}{Effect}{Permanent}{ Changes the shape of your body, reallocating one point from the natural component of AGI to STR every 4 hours and vice versa. Abilities that have STR or AGI requirements that are not met are disabled. It is necessary for the amount of flesh for any weight increase to be available.}

% \spell{Face Sculpting}{4 Hunger, 1 hour}{Biomancy 4}{Medicine 5}{10}{Melee}{Effect}{Permanent}{ Makes a sorcery test of DC 16 to change the appearance of your face. Looks exactly as the caster desires on a critical success and gains the Beautiful or Ugly trait, if desired. Has only minor imperfections on a success. On a graze, has a 50\% chance of becoming ugly. On a failure, becomes ugly. }



% \spell{Healing Leaves}{5 AP + 5 STA}{ 4}{Medicine I}{16}{self, radius of 9m}{Effect}{1 minute}{Generates 5 healing energy leaves that fly around the area. Whenever someone in the area takes damage, a leaf merges with the wound, stopping POD/2 bleeding.}

% \spell{Control Tree }{4AP+1STA}{Biomancy 3}{}{8}{15m}{Effect}{Instantaneous}{ Controls a tree that has a movement mechanism to attack something or move. }

% \spell{Change Voice}{2 AP, 2 STA}{1}{Medicine 1}{14}{Self}{Effect}{Permanent}{Copies the voice of another person that you have heard before.}





\section{Shamanism}

The ability to interpret, negotiate and channel the power of the spiritual forces of the world. This kind of magic is made by communing with spirits, learning about their motivations and convincing them to help somehow.

Spirits make many demands in order to help or lend their abilities. Types of demand: behavioral, offering, task. 

\begin{itemize}
    \item Behaviorial: can be avoid eating meat, wearing something, being honest, etc. Faulty behavior will cease the concession of power. It may take a while for a spirit to be convinced to lend its power.
    \item Offering: can be an item to be placed somewhere, something to be burned, blood to be given. This works like an exchange: give something and get something in return.
    \item Task: can be to defeat someone or drive away something, go somewhere and plant seeds, or any other mission. A task may have a deadline or a dealbreaker that will break the contract.
\end{itemize}

Spirits can be found through synesthesia or through attempts to interact with magical objects. Some spirits can speak, others cannot. Discovering a spirit's intentions can be half the challenge. 

% Spirits can have their power resonate through special items.

There are 5 types of entity that can be negotiated with:

\textbf{ Incorporeal entities:} Shadows, Echoes, wraiths. These have skills and knowledge that they can share. Animancy is the 

\textbf{ Sentient Objects:} work like gatekeepers for magical object, where negotiation is necessary for usage. 

\textbf{ Elementals: } elements of the world that respond to magic. They can be manipulated using alchemy when convinced to help. 

\textbf{ Divinities: } the gods of the world. They respond to clerics and use divination as the skill to use their abilities. 

\textbf{ Life forms: } some living beings, like trees, fungi or bacteria may have a spiritual presence that allows bargaining with. Once the bargain is struck, use Biomancy as the skill to control use the spirit's abilities.

The following spells can only be cast when some sort of deal with a spirit is struck. 

\spell{Aurora }{5 AP + 2 STA or 1 exhaustion for 10 min}{Shamanism 4}{}{10}{5km}{Environmental}{sustained}{Draws or writes a simple and blurred message in the sky that can be seen at night within the spell's range.}

\spell{Call Lightning}{5 AP + 2 STA}{Shamanism 4}{Shamanism 5}{10}{directly from 100m xMM}{Projectile}{instantaneous}{Summons a lightning bolt to a specific target within the electrical storm as a directed effect. The lightning causes 25xDM electric damage. There is no distance penalty.}

\spell{Storm of Lightning}{5-minute ritual, 2 exhaustion}{Shamanism 5}{}{15}{100m, 15m radius on the ground}{Environmental}{5 turns}{As long as there are enough clouds, calls 2 lightning bolts per round in a designated area to hit random targets as a terrain effect. The lightning causes 25xDM electric damage and is defended by reflexes against DL 7.}

\spell{Freeze Water}{5 AP + 3 STA}{Shamanism 3}{}{8}{15m, area of 15 spaces}{Environmental}{1 turn delay, 5 turns}{Freezes water in an area after one turn. A puddle becomes slippery, turning into difficult terrain with a DC of 12. A layer of ice 10 cm thick is created on a water surface, but additional layers can be added if there is enough water. The temperature of the ice increases due to its freezing, causing it to start melting immediately.}

\spell{Create Fog}{8 AP + 3 STA}{Shamanism 2}{}{5}{15m, area with a radius of 50m}{Environmental}{1 minute}{Condenses water in the air over a large area, creating a thick fog that hinders visibility. Cannot be cast in low humidity areas.}

\spell{Rain}{5-minute ritual + 4 Hunger}{Shamanism 4}{}{10}{100m, area with a radius of 45m}{Environmental}{10 minutes}{If there are enough clouds, converts them to a location and makes it rain there, soaking everything. Criticals can cause hail to fall, dealing 15 PEN of physical damage per turn. }

\spell{Landslide}{8 AP + 3 STA}{Shamanism 3}{}{8}{100m, area with a radius of 5m}{Environmental}{1 turn delay}{ Heats and cools a solid, causing expansion and contraction in adjacent planes of the solid, causing it to break. This technique works well on crystalline structures, rocks, and especially wet soil. One turn later, it triggers a landslide or snow that damages anyone still in the area.}

\spell{Strong Wind}{4 AP + 3 STA}{Shamanism 2}{}{10}{self, radius of 9m, length of 30m in the area}{Environmental}{5 turns, 1 turn delay}{ Produces an entropic field that facilitates the movement of wind in one direction, causing a strong gust. Any projectiles going against the wind or perpendicular to it receive -2 to accuracy. Dispels fog and dust.}

% \spell{Call Wave}{5 AP + 6 STA}{ 3}{}{18}{60m, semicircle with a radius of 9/12m}{Explosion}{1 turn delay}{Pulls a wave from a body of water. It hits all who remain in the area and attempts to pull them into the water with a pulling action using the spell's DC instead of the character's Strength.}


\end{multicols}
