\chapter{War}

\begin{multicols}{2}

\section{Character and Motivation}


\section{Encounters and Scenes}

This section includes examples of encounters and their rewards:

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item Assassinate
    \item Steal/Robbery
    \item Sabotage
    \item Spy
    \item Escape
    \item Escort
    \item Deliver
    \item Defend
    \item Rescue
    \item Survive surprise attack
    \item Find traitor
\end{itemize}

\section{War Skills and Tools}

This section includes skills, items, and spells relevant to warfare.

\textbf{Battlefield transformation:} Trenches, hideouts, walls, traps.

\section{Organization of an Army}

\textbf{Rank and reputation:} Rank and reputation play important roles in military hierarchy and command structure.

This section includes various examples of armies and their organization.

\section{Battle}

\textbf{Troops:} Troops are divided into squads. Types of troops have different characteristics: power, consumption, load, and loyalty.

\textbf{Consumption:} Consumption is based on the number of troops.

\textbf{Supplies:} Supply management includes:

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item Supply line
    \item Carrying capacity
    \item Morale
    \item Movement speed
    \item Equipment and preparation: multiplies the power of the troops
\end{itemize}

\textbf{Victory Condition:} A battle ends when:

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item Morale = 0, or
    \item Troops = 0
\end{itemize}

\section{Squads}

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item Treat many soldiers as one unit
    \item Occupies a number of spaces equal to the number of soldiers
    \item Deals attacks based on the number of soldiers in contact
    \item Roll charisma + squad bonus instead of cunning to determine turn order
\end{itemize}

\section{For D\&D}

\enlargethispage{2\baselineskip}

Troops have attacks (damage, range), have defenses (AC, reflexes, will, fortitude as target number), and movement speed.

Life of armies uses a dice pool system. Each damage received causes a loss of dice in the pool.

Confirmed criticals kill leadership.

\textbf{Leadership:}

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item Increases attack dice bonuses and the target number of defenses and troop speed
    \item Grants special attacks for troops
\end{itemize}

\textbf{Terrain/fortifications:}

\begin{itemize}
    \setlength\itemsep{0em}
    \setlength{\parskip}{0em}
    \item Alters speed, defenses, and attack of troops
    \item Allows the use of more dice in attack or defense
    \item Grants special attacks (Cannons, boiling oil)
\end{itemize}

War takes place in turns. Each troop has 1 action per turn. Each turn is a scene.

War has victory conditions for both sides: Kill leadership, conquer territory, defend areas, destroy items, etc.

\end{multicols}